### 2023-03-19 code/analysis_code/exploratory_analysis.qmd
Project Part 3

---
title: "Exploratory analysis"
author: "Irene Cavros and Chris Okitondo"
date: "02/24/2023"
output: html_document
---

This Quarto file loads the cleaned data and does some exploring.

As part of the exploratory analysis, we have produced plots, tables, and other summary quantities for some interesting aspects of our data.

# Setup

```{r, warning=FALSE, message=FALSE}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(ggplot2)
library(haven)
library("kableExtra") # For producing a nice table 1
```

### Load the data

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed_data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

### Double check that we are still ignoring the NA values in the dataset

```{r}
mydata2 <- na.omit(mydata)
```

### Print data

```{r}
print(mydata2)
```

# Data exploration through tables

Showing a bit of code to produce and save a summary table.

This uses a bit of code trickery to get things into the shape of a data frame (for easier saving/showing in manuscript)

```{r}
# looking at the data
summary_df <- dplyr::glimpse(mydata2)
# save to file
summarytable_file = here("results", "summarytable.rds")
saveRDS(summary_df, file = summarytable_file)
```
### Checking names of data
```{r}
names(mydata2)
```

## Data exploration through figures

###Box plots for the flu vaccine uptake and age

```{r}
p1 <- ggplot(mydata2, aes(x=flu_vaccine, y=age)) + 
  geom_boxplot(fill = "gray") +
  labs(title = "Distribution of Age per flu vaccine uptake", x = "Uptake of Flu Vaccine in the past 12 months", y = "Age") +  theme_classic()

p1
```

```{r}
figure_file = here("results","age & flu uptake distribution.png")
ggsave(filename = figure_file, plot=p1) 
```

### Barplot for flu vaccine uptake: 1 for no and 2 for yes

```{r}
ggplot(mydata2, aes(flu_vaccine)) + 
  geom_bar(width=0.7, fill="steelblue") + scale_x_discrete(labels=c('Did not take flu vaccine', 'Did take flu vaccine')) + theme_classic()
```

### Flu vaccine by sex: 1 for no and 2 for yes
```{r}
ggplot(mydata2, aes(x= flu_vaccine, fill = sex)) + 
  geom_bar(position = "dodge") + scale_x_discrete(labels=c('Did not take flu vaccine', 'Did take flu vaccine')) + scale_fill_discrete(labels=c('Male', 'Female')) +  theme_classic()

```

### Flu vaccine by smoking: 1 for nonsmoker, 2 for former smoker, 3 for current smoker

```{r, message=FALSE, warning=FALSE}
ggplot(mydata2, aes(x = flu_vaccine, fill = smoking_status)) +
  geom_histogram(stat="count") + scale_x_discrete(labels=c('Did not take flu vaccine', 'Did take flu vaccine')) + scale_fill_discrete(labels=c('Nonsmoker', 'Former Smoker', 'Current smoker')) +
  ggtitle("Distribution of flu vaccine uptake and smoking status") + theme_classic()
```

### Flu vaccine by health insurance status

```{r, message=FALSE, warning=FALSE}
ggplot(mydata2, aes(x = flu_vaccine, fill = health_ins_status)) +
  geom_histogram(stat="count") + scale_x_discrete(labels=c('Did not take flu vaccine', 'Did take flu vaccine')) + scale_fill_discrete(labels=c('Does not have health insurance', 'Does have health insurance')) +
  ggtitle("Distribution of flu vaccine uptake by health insurance status") + theme_classic()
```

## Tables for patients characteristics

**Table 1. Demographic Characteristics**

```{r, warning=FALSE, message=FALSE}
library(table1)
(table1(~ factor(sex) + age + factor(race) + factor(hispanic) | flu_vaccine, data=mydata2))
```

**Table 2. Socioeconomic Factors**

```{r, warning=FALSE, message=FALSE}
library(table1)
(table1(~ factor(education) + factor(fam_income) | flu_vaccine, data=mydata2))
```

**Table 3. Employment factors**

```{r, warning=FALSE, message=FALSE}
library(table1)
(table1(~ factor(employ_status) + factor(paid_sick) | flu_vaccine, data=mydata2))
```

**Table 4. Healthcare access**

```{r, warning=FALSE, message=FALSE}
library(table1)
(table1(~ factor(health_ins_status) + factor(med_access) + factor(med_care_delayed_cost) | flu_vaccine, data=mydata2))
```

**Table 5. Health status factors**

```{r, warning=FALSE, message=FALSE}
library(table1)
(table1(~ factor(health_status) + factor(bmi) + factor(asthma) + factor(cancer) + factor(chd) + factor(alcohol_status) + factor(smoking_status) | flu_vaccine, data=mydata2))
```

**Table 6. Mental health factors**

```{r, warning=FALSE, message=FALSE}
library(table1)
(table1(~factor(worry_freq) + factor(dep_freq) | flu_vaccine, data=mydata2))

```

## Full table for demogrophics characteristics
```{r, warning=FALSE, message=FALSE}
library(table1)
table_dems <- (table1(~ factor(sex) + age + factor(race) + factor(hispanic) + factor(education) + factor(fam_income) + factor(employ_status) + factor(paid_sick) + factor(health_ins_status) + factor(med_access) + factor(med_care_delayed_cost) + factor(health_status) + factor(bmi) + factor(asthma) + factor(cancer) + factor(chd) + factor(alcohol_status) + factor(smoking_status)| flu_vaccine, data=mydata2))

table_dems

```

### Making demographics table look nicer
```{r}
table_dems_v2 <- t1kable(table_dems)
table_dems_v2
```

### Saving table 1 in the results section
```{r}
saveRDS(table_dems_v2, file = (here("results", "table1_demographics.rds")))
```

### 2023-03-19 code/processing_code/processingfile.qmd
Project Part 3

---
title: "Data Cleaning"
author: "Chris Okitondo & Irene Cavros"
date: "2023-02-20"
output: html_document
editor: 
  markdown: 
    wrap: 72
---

### Setup

*Load needed packages. Don't forget to make sure they are installed. If
not, please install them before moving forward in attempting to run any
code.*

```{r, warning=FALSE, message=FALSE}
library(readxl) #for loading Excel files
library(writexl) #for writing Excel files
library(haven) # for loading SPSS file
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths
library(tidyverse)
```

### Data loading

```{r}
data_location <- here::here("data","raw_data","nhis_00015.sav")
rawdata <- haven:: read_sav(data_location)
data1<- rawdata
```

### Check data

*We are checking the data using several methods.*

```{r}
dplyr::glimpse(data1)
summary(data1)
head(data1)
skimr::skim(data1)
lapply(data1, levels)
```

### Observations for each year: 2012-2021

```{r}
table(data1$YEAR)
```

### Variables for Analysis

*Below is a summary of all of the variables in our dataset.*

-   Demographics: age, sex, race, ethnicity, veteran status

-   Socioeconomic status: education level, total combined family income,
    employment status, paid sick leave as part of job benefits, ratio of
    family income to poverty threshold

-   General health: perceived health status, categorical BMI Conditions:
    ever told had asthma, ever told had cancer, ever told had coronary
    heart disease

-   Health behaviors: alcohol drinking status, cigarette smoking status
    Access to care: has usual place for medical care, medical care
    delayed due to cost (past 12 months)

-   Health insurance: health insurance coverage status Mental health:
    anxiety level, depression level

-   Flu vaccine: Had any flu vaccine, past 12 months. **THIS IS OUR
    OUTCOME OF INTEREST!**

### Re-categorizing variables

*Here we will create the appropriate levels as we re-categorize
variables.* *We will later rename some of the factor levels so that we
do not have to remember or keep referring back to a data dictionary.*

```{r}
# Age
data1$AGE[data1$AGE == 997 | data1$AGE == 998 | data1$AGE == 999] <- NA

# Sex: 1 for Male and 2 for female, 7 or 9 for unknown
data1$SEX[data1$SEX == 7 |  data1$SEX == 9] <- NA

# Race: White:1, Black:2, American Indiana/Alaska Native:3, Asian: 4
data1$RACENEW[data1$RACENEW == 100] <- 1
data1$RACENEW[data1$RACENEW == 200] <- 2
data1$RACENEW[data1$RACENEW == 300] <- 3
data1$RACENEW[data1$RACENEW == 400] <- 4
data1$RACENEW[data1$RACENEW > 400] <- NA

# Hispanic: 1 for no, 2 for yes.
data1$HISPYN[data1$HISPYN == 7 | data1$HISPYN == 8 | data1$HISPYN == 9] <- NA

# Veteran status: 1 for no. 2 for yes
data1$ARMFEV[data1$ARMFEV==10|data1$ARMFEV == 11 |data1$ARMFEV == 12 ] <- 1
data1$ARMFEV[data1$ARMFEV == 20]  <- 2
data1$ARMFEV[data1$ARMFEV==0|data1$ARMFEV==97|data1$ARMFEV ==98|data1$ARMFEV == 99 ]<-NA

# EDUCATION: 1 for less than HS, 2 for HS, 3 for Some college, 4 for bachelor, 5 for graduate degree
data1$EDUC[data1$EDUC >= 100  & data1$EDUC <= 116] <- 1
data1$EDUC[data1$EDUC >= 200  & data1$EDUC <= 202] <- 2
data1$EDUC[data1$EDUC >= 300  & data1$EDUC <= 303] <- 3
data1$EDUC[data1$EDUC == 400] <- 4
data1$EDUC[data1$EDUC >= 500 & data1$EDUC <= 505] <- 5
data1$EDUC[data1$EDUC == 0 | data1$EDUC > 505] <- NA

# EMPLOYMENT STATUS: 1 for not employed, 2 for employed, and 3 for not in labor force
data1$EMPSTAT[data1$EMPSTAT == 200] <- 1
data1$EMPSTAT[data1$EMPSTAT == 100  |  data1$EMPSTAT == 111 | data1$EMPSTAT == 112 | data1$EMPSTAT == 120] <- 2
data1$EMPSTAT[data1$EMPSTAT ==220] <- 3
data1$EMPSTAT[data1$EMPSTAT == 0  | data1$EMPSTAT ==900 | data1$EMPSTAT ==997 | data1$EMPSTAT ==998 | data1$EMPSTAT ==999] <- NA

# PAID Sick: 1 for no, 2 for yes (0,9,8,7)
data1$PAIDSICK[data1$PAIDSICK == 220] <- 3
data1$PAIDSICK[data1$PAIDSICK == 0 | data1$PAIDSICK == 7 | data1$PAIDSICK == 8 | data1$PAIDSICK == 9] <- NA

# INCOME: 0 - 49,999: 1, 50,000-99,999: 2, AND +100k: 3
data1$INCFAM07ON[data1$INCFAM07ON >= 10  & data1$INCFAM07ON <= 12] <- 1
data1$INCFAM07ON[data1$INCFAM07ON >= 22  & data1$INCFAM07ON <= 23] <- 2
data1$INCFAM07ON[data1$INCFAM07ON == 24] <- 3
data1$INCFAM07ON[data1$INCFAM07ON > 24] <- NA
data1$INCFAM07ON[data1$INCFAM07ON == 20 | data1$INCFAM07ON == 21] <- NA

# HEALTH STATUS: 1 Excellent, 2: Very good, 3: Good, 4: Fair, 5: Poor
data1$HEALTH [data1$HEALTH  > 5] <- NA

# BMI CAT: 1 for BMI < 18.5 (Underweight), 2 for 18.5 - 25 (Normalt), 3 for 25 - 30 (Overweight), 4 for >30 (Obese)
data1$BMICAT [data1$BMICAT  == 0 | data1$BMICAT  == 9] <- NA

# USUAL PLACE FOR MEDICAL CARE: 1 for no, 2 and 3 for yes
data1$USUALPL[data1$USUALPL == 1] <- 1
data1$USUALPL[data1$USUALPL == 2 | data1$USUALPL == 3] <- 2
data1$USUALPL[data1$USUALPL==0 | data1$USUALPL==7| data1$USUALPL==8| data1$USUALPL==9] <- NA

# Medical care delayed due to cost: 1 for no, 2 for yes
data1$DELAYCOST[data1$DELAYCOST ==0 | data1$DELAYCOST ==7| data1$DELAYCOST ==8| data1$DELAYCOST==9] <- NA

# Health insurance coverage: 1 for no, 2 for yes
data1$HINOTCOVE[data1$HINOTCOVE == 2] <- 3
data1$HINOTCOVE[data1$HINOTCOVE == 1] <- 2
data1$HINOTCOVE[data1$HINOTCOVE == 3] <- 1
data1$HINOTCOVE[data1$HINOTCOVE == 9] <- NA

# Had flu vaccine, past 12 months
data1$VACFLU12M [data1$VACFLU12M  == 0|data1$VACFLU12M  == 3 |data1$VACFLU12M ==7|data1$VACFLU12M ==8|data1$VACFLU12M ==9] <- NA

# Had covid, according to test
data1$CVDTESTRSLT [data1$CVDTESTRSLT  == 0|data1$CVDTESTRSLT  == 3 |data1$CVDTESTRSLT ==7|data1$CVDTESTRSLT ==8|data1$CVDTESTRSLT ==9] <- NA

# Ever told has Asthma: ASTHMAEV: 1 for no, 2 for yes
data1$ASTHMAEV[data1$ASTHMAEV == 0|data1$ASTHMAEV==7|data1$ASTHMAEV==9] <- NA

# Ever told has Cancer: CANCEREV: 1 for no, 2 for yes
data1$CANCEREV[data1$CANCEREV == 0|data1$CANCEREV==7|data1$CANCEREV==9] <- NA

# Ever told has CHD: CHEARTDIEV 1 for no, 2 for yes
data1$CHEARTDIEV[data1$CHEARTDIEV==0|data1$CHEARTDIEV==7|data1$CHEARTDIEV==9] <- NA

# ALCSTAT1: Alcohol drinking: 1 for no drinker, 2 for former drinker, 3 for current drinker
data1$ALCSTAT1 [data1$ALCSTAT1 ==0|data1$ALCSTAT1 ==9] <- NA

# Smoking status: 1 for no smoker, 2 for former smoker, 3 for current smoker
data1$SMOKESTATUS2[data1$SMOKESTATUS2 == 10|data1$SMOKESTATUS2  == 11|data1$SMOKESTATUS2==12] <-3
data1$SMOKESTATUS2 [data1$SMOKESTATUS2 == 20 ] <- 2
data1$SMOKESTATUS2 [data1$SMOKESTATUS2 == 30 ] <- 1
data1$SMOKESTATUS2 [data1$SMOKESTATUS2 > 30 ] <- NA

# Worry, nervous, or anxious frequency: WORFREQ. 1 Daily, 2 weekly, 3 Monthly, 4 a few times a year, 5 Never
data1$WORFREQ [data1$WORFREQ ==0|data1$WORFREQ ==7|data1$WORFREQ ==8|data1$WORFREQ ==9] <- NA

# Depression frequency: DEPFREQ. 1 Daily, 2 weekly, 3 Monthly, 4 a few times a year, 5 Never
data1$DEPFREQ[data1$DEPFREQ==0|data1$DEPFREQ==7|data1$DEPFREQ==8|data1$DEPFREQ==9] <- NA
```

*Let's check to confirm all of our re-categorizations look right.
```{r}
summary(data1)
```

### Renaming variables

```{r}
data1 <- rename(data1, age = AGE, sex = SEX, race = RACENEW, hispanic = HISPYN, veteran_status = ARMFEV, education = EDUC, employ_status = EMPSTAT, paid_sick = PAIDSICK, fam_income = INCFAM07ON, pov_ratio = POVERTY, health_status = HEALTH, bmi = BMICAT, med_access = USUALPL, med_care_delayed_cost = DELAYCOST, health_ins_status = HINOTCOVE, flu_vaccine = VACFLU12M, asthma = ASTHMAEV, cancer = CANCEREV, chd = CHEARTDIEV, alcohol_status = ALCSTAT1, smoking_status = SMOKESTATUS2, worry_freq	= WORFREQ, dep_freq = DEPFREQ, covid = CVDTESTRSLT)
```

### Check new variable names

```{r}
names(data1)
```

### Converting some variables to characters

```{r}
data1 <- data1 %>% 
     mutate_at(c(8:30), as.character)
```

### Renaming factor levels
```{r}
#recode sex variable
data1 <- data1 %>% 
     mutate(sex = recode_factor(data1$sex, `1` = "male", `2` = "female"))

#recode race variable
data1 <- data1 %>% 
     mutate(race = recode_factor(data1$race, `1` = "white", `2` = "black", `3` = "native american", `4` = "asian"))

#recode hispanic variable
data1 <- data1 %>% 
     mutate(hispanic = recode_factor(data1$hispanic, `1` = "non-hispanic", `2` = "hispanic"))

#recode veteran_status variable
data1 <- data1 %>% 
     mutate(veteran_status = recode_factor(data1$veteran_status, `1` = "non-veteran", `2` = "veteran"))

#recode education variable
data1 <- data1 %>% 
     mutate(education = recode_factor(data1$education, `1` = "less_than_hs", `2` = "high_school", `3` = "some_college", `4` = "bachelor", `5` = "graduate"))

#recode employment status variable
data1 <- data1 %>% 
     mutate(employ_status = recode_factor(data1$employ_status, `1` = "unemployed", `2` = "employed", `3` = "retired"))

#recode paid sick leave variable
data1 <- data1 %>% 
     mutate(paid_sick = recode_factor(data1$paid_sick, `1` = "no", `2` = "yes"))

#recode household income variable
data1 <- data1 %>% 
     mutate(fam_income = recode_factor(data1$fam_income, `1` = "< $50,000", `2` = "$50,000-$99,999", `3` = "$100,000+"))

#recode health status variable
data1 <- data1 %>% 
     mutate(health_status = recode_factor(data1$health_status, `1` = "excellent", `2` = "very_good", `3` = "good", `4` = "fair", `5` = "poor"))

#recode bmi variable
data1 <- data1 %>% 
     mutate(bmi = recode_factor(data1$bmi, `1` = "underweight", `2` = "average", `3` = "overweight", `4` = "obese"))

#recode variable on usual place for accessing medical care 
data1 <- data1 %>% 
     mutate(med_access = recode_factor(data1$med_access, `1` = "no", `2` = "yes"))

#recode variable on medical care delayed due to cost
data1 <- data1 %>% 
     mutate(med_care_delayed_cost = recode_factor(data1$med_care_delayed_cost, `1` = "no", `2` = "yes"))

#recode health insurance coverage variable
data1 <- data1 %>% 
     mutate(health_ins_status = recode_factor(data1$health_ins_status, `1` = "no", `2` = "yes"))

#recode flu vaccine variable
data1 <- data1 %>% 
     mutate(flu_vaccine = recode_factor(data1$flu_vaccine, `1` = "no", `2` = "yes"))

#recode COVID variable
data1 <- data1 %>% 
     mutate(covid = recode_factor(data1$covid, `1` = "no", `2` = "yes"))

#recode asthma variable
data1 <- data1 %>% 
     mutate(asthma = recode_factor(data1$asthma, `1` = "no", `2` = "yes"))

#recode cancer variable
data1 <- data1 %>% 
     mutate(cancer = recode_factor(data1$cancer, `1` = "no", `2` = "yes"))

#recode heart disease variable
data1 <- data1 %>% 
     mutate(chd = recode_factor(data1$chd, `1` = "no", `2` = "yes"))

#recode alcohol variable
data1 <- data1 %>% 
     mutate(alcohol_status = recode_factor(data1$alcohol_status, `1` = "nondrinker", `2` = "former_drinker", `3` = "drinker"))

#recode smoking variable
data1 <- data1 %>% 
     mutate(smoking_status = recode_factor(data1$smoking_status, `1` = "nonsmoker", `2` = "former_smoker", `3` = "smoker"))

#recode worry frequency variable
data1 <- data1 %>% 
     mutate(worry_freq = recode_factor(data1$worry_freq, `1` = "daily", `2` = "weekly", `3` = "monthly", `4` = "few_per_year", `5` = "never"))

#recode depression frequency variable
data1 <- data1 %>% 
     mutate(dep_freq = recode_factor(data1$dep_freq, `1` = "daily", `2` = "weekly", `3` = "monthly", `4` = "few_per_year", `5` = "never"))
```

### Creating a new dataframe without the COVID variable

*Because we only have COVID-19 information for 2020 and 2021, we will
omit that variable for now and probably use it for sensitivity analysis.*

```{r}
data1 <- data1[,-30]
```

### Check all factors and variables

*Now is a good time to ensure you do not have any unexpected values in
the factors or variables.*

```{r}
View(data1)
```

### Creating a new file for processed data

```{r}
processeddata <- data1
```

### Saving the processed data

```{r}
save_data_location <- here::here("data","processed_data","processeddata.rds")
saveRDS(processeddata, file = save_data_location)
```

```{r}
unlink(".RData")
```

### 2023-03-19 products/manuscript/Manuscript.qmd
Project Part 3

---
title: "Which groups are at highest risk for low uptake of flu vaccine?"
author: Chris Okitondo and Irene Cavros
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
editor: 
  markdown: 
    wrap: 72
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

# Summary/Abstract

[\*\*To be completed (PART 5)]{.underline}

# Introduction

### General Background Information

[\*\*To be completed (PART 4)]{.underline}

### Questions/Hypotheses to be addressed

Using the aforementioned database, our research question of interest is
as follows:

Which groups are at highest risk for low uptake of flu vaccine?

# Methods

### Description of data and data source

The data source used for this project is the National Health Interview
Survey (NHIS.) It is a harmonized set of data beginning in 1963 with
information on general health status, acute and chronic illness,
functional limitations, access to care, insurance coverage, and health
behaviors for the U.S. population. On average, the survey covers 100,000
persons in 45,000 households each year. PUMS NHIS currently contains
over 17,000 integrated variables from 1963 to present.

As the outcome of interest, researchers focused on those who, in this
household survey between the years of 2012 and 2021, answered no when
they were asked whether they had ANY flu vaccine in the past 12 months.

Variables assessed are as follows:

1\. Demographics: age, sex, race, ethnicity, veteran status

2\. Socioeconomic status: education level, total combined family income,
employment status, paid sick leave as part of job benefits, ratio of
family income to poverty threshold

3\. General health: perceived health status, categorical BMI

4\. Conditions: ever told had asthma, ever told had cancer, ever told
had coronary heart disease

5\. Health behaviors: alcohol drinking status, cigarette smoking status

6\. Access to care: has usual place for medical care, medical care
delayed due to cost (past 12 months)

7\. Health insurance: health insurance coverage status

8\. Mental health: anxiety level, depression level

#### Study population

The raw data for our period of interest directly extracted from the NHIS
IPUMS survey initially had had 792,916 observations. After processing
the raw data and removing all observations with missing data in any of
our predictors of interest, 106980 individuals were included in our
analysis. Table 1 below provides a summary of demographic
characteristics of our study population.

[\*\*To be completed (PART 4)]{.underline}

### Statistical analysis

*Explain anything related to your statistical analyses.*

[\*\*To be completed (PART 4)]{.underline}

# Results

### Exploratory/Descriptive analysis

\*\* Table 1: Demographic characteristics of the study population \*\*

### Loading Table 1 from the Exploratory Analysis

```{r, warning=FALSE, message=FALSE}
result_table1 <- readRDS(here("results", "table1_demographics.rds"))
result_table1
```

\*\*

### Full analysis

##### Model 1: Focusing on only demogrphics characteristics using the flu vaccine uptake as our outcome

```{r, warning=FALSE, message=FALSE}
result_table2 <- readRDS(here("results", "demographics_model.rds"))
result_table2
```

### Model 2: Using all our predictors of interest in the model with flu vaccine uptake as our outcome

```{r, warning=FALSE}
result_table3 <- readRDS(here("results", "full_model.rds"))
result_table3
```

\*\*

# Discussion

### Summary and Interpretation

*Summarize what you did, what you found and what it means.*

[\*\*To be completed (PART 5)]{.underline}

### Strengths and Limitations

*Discuss what you perceive as strengths and limitations of your
analysis.*

[\*\*To be completed (PART 5)]{.underline}

### Conclusions

*What are the main take-home messages?*

[\*\*To be completed (PART 5)]{.underline}

# References

*Include citations in your Rmd file using bibtex, the list of references
will automatically be placed at the end*

[\*\*To be completed (PART 5)]{.underline}

### 2023-04-07 products/manuscript/Manuscript.qmd
Project Part 4 v1

@@
 
-[\*\*To be completed (PART 5)]{.underline}
+Influenza, or flu, is an extremely contagious respiratory illness caused
+by a virus. It can cause a range of symptoms, including fever, cough,
+sore throat, body aches, headache, chills, and fatigue (CDC, Key Facts
+about Seasonal Flu Vaccine, 2022). It can be particularly dangerous for
+certain groups, such as older adults, young children, and individuals
+with underlying health issues, often resulting in severe illness or even
+death  Although influenza vaccinations are widely accessible, many
+people in the United States do not receive them as recommended. This
+study examines influenza vaccination uptake among US adults from
+2012-2021. Using a subset of the National Health Interview Survey (NHIS)
+data set, researchers provide important insights into which groups are
+at highest risk for under-coverage of flu vaccine in order to better
+inform targeted public health interventions for improving uptake.
 
@@
 
-[\*\*To be completed (PART 4)]{.underline}
+Influenza, commonly known as the flu, is a contagious respiratory
+illness which comes from influenza viruses. It can cause mild to severe
+illness, and in some cases, it can lead to hospitalization or even
+death. Symptoms of the flu include fever, cough, sore throat, body
+aches, headache, chills, and fatigue (CDC, About Flu, 2022). It can
+spread easily from person to person through droplets that are released
+when infected individuals cough, sneeze, or talk.
 
-### Questions/Hypotheses to be addressed
+Vaccines are a crucial tool in preventing the spread of infectious
+diseases, including the flu. On an individual level, the best way to
+protect against the flu is by getting vaccinated each year. At the
+population level, high vaccine coverage is a powerful strategy for
+reducing spread of the virus (CDC, Key Facts about Seasonal Flu Vaccine,
+2022).
 
-Using the aforementioned database, our research question of interest is
-as follows:
+The U.S. Centers for Disease Control and Prevention (CDC) recommends
+that everyone six months of age and older receive a flu vaccine each
+year, with rare exceptions (CDC, Key Facts about Seasonal Flu Vaccine,
+2022). Influenza is a constantly evolving virus that mutates rapidly,
+and uncontrolled spread gives rise to many different strains, so new
+vaccines are developed each year to protect against the most common
+strains of the virus that are expected to circulate during the upcoming
+flu season. It is especially important for individuals who are at higher
+risk of severe illness or complications from the flu, such as young
+children, older adults, pregnant women, and individuals with underlying
+medical conditions.
 
-Which groups are at highest risk for low uptake of flu vaccine?
+In terms of the availability and cost, there have been significant
+changes for flu vaccines over the past 50 years (CDC, Influenza Historic
+Timeline, 2019). In the 1960s, flu vaccines were not widely available
+and were primarily given to people considered to be at high risk for
+complications. The cost of the vaccine was relatively high and was not
+covered by most insurance plans. In the 1970s and 1980s, flu vaccines
+became more widely available and were recommended for more people,
+including children and healthy adults. The cost of the vaccine began to
+decrease as more people received it, and insurance coverage for the
+vaccine became more common. In the 1990s and 2000s, the availability of
+flu vaccines continued to increase, and the cost of the vaccine became
+more affordable for a larger proportion of the U.S. population.
+
+Today, flu vaccines are frequently available in doctor's offices,
+pharmacies, community clinics, workplace clinics, school-based clinics,
+and local health departments, among others. They are often covered by
+insurance plans, which has increased their accessibility. Many public
+health clinics and community organizations also offer flu vaccines at
+low or no cost, in order to better target historically unreached
+populations.
+
+Despite the availability and recommendation of influenza vaccines, many
+people in the United States still do not receive them as recommended. In
+fact, the CDC reported that only 49.2% of adults aged 18 years and older
+received a flu vaccine during the 2020-2021 flu season (CDC, Inequities
+in Flu Vaccine Uptake 2022). This low uptake of influenza vaccine
+highlights the need for understanding the factors that contribute to
+under-coverage of the vaccine in certain populations.
+
+Racial/ethnic minorities have a long history of reduced vaccination
+rates relative to white Americans (Khan 2018). Since 2010, flu
+vaccination coverage has been consistently lower among Black, Hispanic,
+and AI/AN adults. During the 2021--2022 season, flu vaccination coverage
+was 54% among White adults, 42% among Black adults, 38% among Hispanic
+adults, and 41% among AI/AN adults (CDC, Inequities in Flu Vaccine
+Uptake 2022). There are many reasons for these inequities, including
+lack of access to health care and insurance, missed opportunities to
+vaccinate, and misinformation and distrust.
+
+Racial and ethnic minorities have a longstanding history of lower
+vaccination rates compared to white Americans. Since 2010, Black,
+Hispanic, and American Indian/Alaska Native adults have consistently had
+lower flu vaccination rates than White adults. In the 2021-2022 flu
+season, vaccination coverage for White adults was 54%, while it was 42%
+for Black adults, 38% for Hispanic adults, and 41% for AI/AN adults
+(CDC, Inequities in Flu Vaccine Uptake 2022).  The reasons for these
+disparities are multifactorial, and may include limited access to
+healthcare and insurance, missed opportunities for vaccination, and
+misinformation and mistrust.
+
+In prior studies, lower income levels were associated with decreased
+odds of influenza vaccine receipt (Gaskin 2022, Vashist 2022), and
+having health insurance coverage has also been shown to be associated
+with increased influenza vaccine coverage (Cambou 2021, Vashist 2022).
+Influenza vaccination rates were also found to be lower in rural than in
+urban areas among adults of all age groups and both genders (Jain 2022).
+However, there are still many variables of interest from the NHIS
+dataset that have yet to be evaluated in detail. Additionally, the
+periods of study from all prior research all end prior to the onset of
+the COVID-19 pandemic, which could have altered or heightened
+disparities even more.
 
@@
 The data source used for this project is the National Health Interview
-Survey (NHIS.) It is a harmonized set of data beginning in 1963 with
+Survey (NHIS). It is a harmonized set of data beginning in 1963 with
 information on general health status, acute and chronic illness,
@@
 behaviors for the U.S. population. On average, the survey covers 100,000
-persons in 45,000 households each year. PUMS NHIS currently contains
-over 17,000 integrated variables from 1963 to present.
+persons in 45,000 households each year. It is conducted in-person and by
+telephone, and its sample is designed to be representative of the
+civilian United States population (NHIS, About).
 
-As the outcome of interest, researchers focused on those who, in this
-household survey between the years of 2012 and 2021, answered no when
-they were asked whether they had ANY flu vaccine in the past 12 months.
+#### Study population
 
-Variables assessed are as follows:
+Individuals were eligible for the study if they were interviewed as part
+of the NHIS household survey between the years of 2012 and 2021. Table 1
+below provides a summary of demographic characteristics of our study
+population.
+
+### Statistical analysis
+
+The raw data for the period of interest directly extracted from the NHIS
+IPUMS survey initially had had 792,916 observations. Researchers then
+selected their predictors of interest to model the association between
+the outcome and variables. Variables assessed are as follows:
 
@@
 2\. Socioeconomic status: education level, total combined family income,
-employment status, paid sick leave as part of job benefits, ratio of
-family income to poverty threshold
+employment status, paid sick leave as part of job benefits
 
@@
 
-8\. Mental health: anxiety level, depression level
+8\. Mental health: anxiety level, depression level.
 
-#### Study population
-
-The raw data for our period of interest directly extracted from the NHIS
-IPUMS survey initially had had 792,916 observations. After processing
-the raw data and removing all observations with missing data in any of
-our predictors of interest, 106980 individuals were included in our
-analysis. Table 1 below provides a summary of demographic
-characteristics of our study population.
-
-[\*\*To be completed (PART 4)]{.underline}
-
-### Statistical analysis
-
-*Explain anything related to your statistical analyses.*
-
-[\*\*To be completed (PART 4)]{.underline}
+\
+After processing the raw data and removing all observations with missing
+data for any of the predictors of interest, the remaining 106,980
+individuals were included in the analysis. Because of the binary
+outcome, logistic regression was utilized to model the association
+between the outcome and variables of interest. Researchers utilized
+machine learning to split data and create a workflow. Data was split
+randomly where 75% of included participants made up the training dataset
+and 25% made up the testing dataset.
 

### 2023-04-07 code/analysis_code/modeleval.qmd
Adding model evaluation

---
title: "Model Evaluation"
author: "Chris Okitondo And Irene Cavros"
date: "04/06/2023"
output: html_document
---

### Loading necessary packages
```{r, message=FALSE, warning=FALSE}
library(here)
library(tidyverse)
library(tidymodels)
library(performance)
library(dplyr)
library(skimr)
```

### Loading previously processed data
```{r}
#Path to data. 
data_location <- here::here("data","processed_data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

### Ensure we are ignoring the NA values in the dataset
```{r}
mydata <- na.omit(mydata)
```

### Data splitting
* Here, we are splitting our data into two subsets. 3/4 for the training and 1/4 for test.
```{r}
# To maintain reproducible results when re-done.
set.seed(1234)

# Proceeding with 75% for training and 25% for test data
split <- initial_split(mydata, prop = 3/4)
```

#### Dataframe for the training and test data
```{r}
train_data <- training(split)
test_data  <- testing(split)
```

* Training data set now contains 80235 observations while test data contains 26745.

### Workflow creation and model fitting

#### FULL MODEL: Creating recipe that fits a logistic model with flu_vaccine as the outcome and all the predictors of interest in the data

```{r, warning=FALSE, message=FALSE}
# Outcome is flu_vaccine. The rest of variables are our predictors of interest
# Recipe #1: flu_vaccine predicted by all predictors of interest
flu_rec <- recipe(flu_vaccine ~ age + sex + race + hispanic + education + fam_income + asthma + cancer + chd + smoking_status+ alcohol_status, data = train_data)

# Model: logistic regression using GLM engine
flu_mod <- logistic_reg() %>% 
                    set_engine("glm")

# Workflow: Pairing model and recipe 
flu_workflow <- workflow() %>% 
  add_model(flu_mod) %>% 
  add_recipe(flu_rec)
# Pring workflow
flu_workflow

# Fitting the model to a training dataset
flu_fit <- 
 flu_workflow %>% 
  fit(data = train_data)

# Looking at the model output
flu_fit %>% 
  extract_fit_parsnip() %>% 
  tidy()
```

#### FULL MODEL: USING trained workflow to predict unseen test data

```{r, warning = FALSE, message=FALSE}
# Using the trained workflow (flu_fit) to predict with the unseen test data
predict(flu_fit, test_data)

# Using argument() with the model plus test data for saving them together
flu_aug <- 
  augment(flu_fit, test_data)

flu_aug %>%
  select(flu_vaccine, .pred_no, .pred_yes)

# Plotting ROC curve
flu_aug %>% 
  roc_curve(truth = flu_vaccine, .pred_no) %>% 
  autoplot()

# Using roc_auc() to get the estimates
flu_aug %>% 
  roc_auc(truth = flu_vaccine, .pred_no)
```

#### ALTERNATIVE MODEL: Outcome is still flu_vaccine and predictor is race 
```{r, warning=FALSE, message=FALSE}
# Using flu_vaccine as a categorical outcome of interest and race as main predictor
flu_rec2 <- recipe(flu_vaccine ~ race, data = train_data)

# Fitting the logistic model
flu_mod2 <- logistic_reg() %>% 
                    set_engine("glm")

# Modelling workflow for pairing model and recipe 
flu_workflow2 <- workflow() %>% 
  add_model(flu_mod2) %>% 
  add_recipe(flu_rec2)

flu_workflow2
# Using the resulting predictors for preparing recipe and training the model
flu_fit2 <- 
 flu_workflow2 %>% 
  fit(data = train_data)

# Pulling the fitted model object and using tidy() function for getting a tidy tibble of model coefficients
flu_fit2 %>% 
  extract_fit_parsnip() %>% 
  tidy()
```

**ALTERNATIVE MODEL: USING TRAINED WORKFLOW TO PREDICT**

```{r, warning = FALSE, message=FALSE}
# Using the trained data to predict with the unseen test data
predict(flu_fit2, test_data)

# Using argument() with the model plus test data for saving them together
flu_aug2 <- 
  augment(flu_fit2, test_data)

flu_aug2 %>%
  select(flu_vaccine, .pred_no, .pred_yes)

# Creating ROC curve and piping to the autoplot() method
flu_aug2 %>% 
  roc_curve(truth = flu_vaccine, .pred_no) %>% 
  autoplot()
# Estimating the area under the curve 
flu_aug2 %>% 
  roc_auc(truth = flu_vaccine, .pred_no)
```

Given this result, the full model using all variables of interest as predictors perform better the reduced model with only race as predictor. The full model as a ROC-AUC of 0.70 compared the reduced model with ROC-AUC of 0.53

### 2023-04-21 products/manuscript/Manuscript.qmd

@@
     highlight-style: github
+bibliography: ../references.bib
+csl: ../apa.csl
 editor: 
@@
 library(knitr)
+library(png)
+
+remotes::install_github("paleolimbot/rbbt") #for integration of BibTex Zotero citations
+library(rbbt)
 ```
@@
 
-Influenza, or flu, is an extremely contagious respiratory illness caused
-by a virus. It can cause a range of symptoms, including fever, cough,
-sore throat, body aches, headache, chills, and fatigue (CDC, Key Facts
-about Seasonal Flu Vaccine, 2022). It can be particularly dangerous for
-certain groups, such as older adults, young children, and individuals
-with underlying health issues, often resulting in severe illness or even
-death  Although influenza vaccinations are widely accessible, many
-people in the United States do not receive them as recommended. This
-study examines influenza vaccination uptake among US adults from
-2012-2021. Using a subset of the National Health Interview Survey (NHIS)
-data set, researchers provide important insights into which groups are
-at highest risk for under-coverage of flu vaccine in order to better
-inform targeted public health interventions for improving uptake.
+Influenza, or flu, is an extremely contagious respiratory illness caused by a virus.
+It can cause a range of symptoms, including fever, cough, sore throat, body aches, headache, chills, and fatigue (@cdc2022c.) It can be particularly dangerous for certain groups, such as older adults, young children, and individuals with underlying health issues, often resulting in severe illness or even death.
+Although influenza vaccinations are widely accessible, many people in the United States do not receive them as recommended.
+This study examines influenza vaccination uptake among US adults from 2012-2021.
+Using a subset of the National Health Interview Survey (NHIS) data set, researchers provide important insights into which groups are at highest risk for under-coverage of flu vaccine in order to better inform targeted public health interventions for improving uptake.
 
@@
 
-### General Background Information
+Influenza, commonly known as the flu, is a contagious respiratory illness which comes from influenza viruses.
+It can cause mild to severe illness, and in some cases, it can lead to hospitalization or even death.
+Symptoms of the flu include fever, cough, sore throat, body
+aches, headache, chills, and fatigue (@cdc2022.) It can spread easily from person to person through droplets that are released infected individuals cough, sneeze, or talk.
 
-Influenza, commonly known as the flu, is a contagious respiratory
-illness which comes from influenza viruses. It can cause mild to severe
-illness, and in some cases, it can lead to hospitalization or even
-death. Symptoms of the flu include fever, cough, sore throat, body
-aches, headache, chills, and fatigue (CDC, About Flu, 2022). It can
-spread easily from person to person through droplets that are released
-when infected individuals cough, sneeze, or talk.
+Vaccines are a crucial tool in preventing the spread of infectious diseases, including the flu.
+On an individual level, the best way to protect against the flu is by getting vaccinated each year.
+At the population level, high vaccine coverage is a powerful strategy for reducing spread of the virus (@cdc2022c.)
 
-Vaccines are a crucial tool in preventing the spread of infectious
-diseases, including the flu. On an individual level, the best way to
-protect against the flu is by getting vaccinated each year. At the
-population level, high vaccine coverage is a powerful strategy for
-reducing spread of the virus (CDC, Key Facts about Seasonal Flu Vaccine,
-2022).
+The U.S.
+Centers for Disease Control and Prevention (CDC) recommends that everyone six months of age and older receive a flu vaccine each year, with rare exceptions (@cdc2022c.) Influenza is a constantly evolving virus that mutates rapidly, and uncontrolled spread gives rise to many different strains, so new vaccines are developed each year to protect against the most common strains of the virus that are expected to circulate during the upcoming flu season.
+It is especially important for individuals who are at higher risk of severe illness or complications from the flu, such as young children, older adults, pregnant women, and individuals with underlying medical conditions.
 
-The U.S. Centers for Disease Control and Prevention (CDC) recommends
-that everyone six months of age and older receive a flu vaccine each
-year, with rare exceptions (CDC, Key Facts about Seasonal Flu Vaccine,
-2022). Influenza is a constantly evolving virus that mutates rapidly,
-and uncontrolled spread gives rise to many different strains, so new
-vaccines are developed each year to protect against the most common
-strains of the virus that are expected to circulate during the upcoming
-flu season. It is especially important for individuals who are at higher
-risk of severe illness or complications from the flu, such as young
-children, older adults, pregnant women, and individuals with underlying
-medical conditions.
+In terms of the availability and cost, there have been significant changes for flu vaccines over the past 50 years (@InfluenzaHistoricTimeline2022.) In the 1960s, flu vaccines were not widely available and were primarily given to people considered to be at high risk for complications.
+The cost of the vaccine was relatively high and was not covered by most insurance plans.
+In the 1970s and 1980s, flu vaccines became more widely available and were recommended for more people, including children and healthy adults.
+The cost of the vaccine began to decrease as more people received it, and insurance coverage for the vaccine became more common.
+In the 1990s and 2000s, the availability of flu vaccines continued to increase, and the cost of the vaccine became more affordable for a larger proportion of the U.S.
+population.
 
-In terms of the availability and cost, there have been significant
-changes for flu vaccines over the past 50 years (CDC, Influenza Historic
-Timeline, 2019). In the 1960s, flu vaccines were not widely available
-and were primarily given to people considered to be at high risk for
-complications. The cost of the vaccine was relatively high and was not
-covered by most insurance plans. In the 1970s and 1980s, flu vaccines
-became more widely available and were recommended for more people,
-including children and healthy adults. The cost of the vaccine began to
-decrease as more people received it, and insurance coverage for the
-vaccine became more common. In the 1990s and 2000s, the availability of
-flu vaccines continued to increase, and the cost of the vaccine became
-more affordable for a larger proportion of the U.S. population.
+Today, flu vaccines are frequently available in doctor's offices, pharmacies, community clinics, workplace clinics, school-based clinics, and local health departments, among others.
+They are often covered by insurance plans, which has increased their accessibility.
+Many public health clinics and community organizations also offer flu vaccines at low or no cost, in order to better target historically unreached populations.
 
-Today, flu vaccines are frequently available in doctor's offices,
-pharmacies, community clinics, workplace clinics, school-based clinics,
-and local health departments, among others. They are often covered by
-insurance plans, which has increased their accessibility. Many public
-health clinics and community organizations also offer flu vaccines at
-low or no cost, in order to better target historically unreached
-populations.
+Despite the availability and recommendation of influenza vaccines, many people in the United States still do not receive them as recommended.
+In fact, the CDC reported that only 49.2% of adults aged 18 years and older received a flu vaccine during the 2020-2021 flu season (@cdc2022b.) This low uptake of influenza vaccine highlights the need for understanding the factors that contribute to under-coverage of the vaccine in certain populations.
 
-Despite the availability and recommendation of influenza vaccines, many
-people in the United States still do not receive them as recommended. In
-fact, the CDC reported that only 49.2% of adults aged 18 years and older
-received a flu vaccine during the 2020-2021 flu season (CDC, Inequities
-in Flu Vaccine Uptake 2022). This low uptake of influenza vaccine
-highlights the need for understanding the factors that contribute to
-under-coverage of the vaccine in certain populations.
+Racial/ethnic minorities have a long history of reduced vaccination rates relative to white Americans (@khan_etal2018.) Since 2010, flu vaccination coverage has been consistently lower among Black, Hispanic, and AI/AN adults.
+During the 2021--2022 season, flu vaccination coverage was 54% among White adults, 42% among Black adults, 38% among Hispanic adults, and 41% among AI/AN adults (@cdc2022b.) There are many reasons for these inequities, including lack of access to health care and insurance, missed opportunities to vaccinate, and misinformation and distrust.
 
-Racial/ethnic minorities have a long history of reduced vaccination
-rates relative to white Americans (Khan 2018). Since 2010, flu
-vaccination coverage has been consistently lower among Black, Hispanic,
-and AI/AN adults. During the 2021--2022 season, flu vaccination coverage
-was 54% among White adults, 42% among Black adults, 38% among Hispanic
-adults, and 41% among AI/AN adults (CDC, Inequities in Flu Vaccine
-Uptake 2022). There are many reasons for these inequities, including
-lack of access to health care and insurance, missed opportunities to
-vaccinate, and misinformation and distrust.
-
-Racial and ethnic minorities have a longstanding history of lower
-vaccination rates compared to white Americans. Since 2010, Black,
-Hispanic, and American Indian/Alaska Native adults have consistently had
-lower flu vaccination rates than White adults. In the 2021-2022 flu
-season, vaccination coverage for White adults was 54%, while it was 42%
-for Black adults, 38% for Hispanic adults, and 41% for AI/AN adults
-(CDC, Inequities in Flu Vaccine Uptake 2022).  The reasons for these
-disparities are multifactorial, and may include limited access to
-healthcare and insurance, missed opportunities for vaccination, and
-misinformation and mistrust.
-
-In prior studies, lower income levels were associated with decreased
-odds of influenza vaccine receipt (Gaskin 2022, Vashist 2022), and
-having health insurance coverage has also been shown to be associated
-with increased influenza vaccine coverage (Cambou 2021, Vashist 2022).
-Influenza vaccination rates were also found to be lower in rural than in
-urban areas among adults of all age groups and both genders (Jain 2022).
-However, there are still many variables of interest from the NHIS
-dataset that have yet to be evaluated in detail. Additionally, the
-periods of study from all prior research all end prior to the onset of
-the COVID-19 pandemic, which could have altered or heightened
-disparities even more.
+In prior studies, lower income levels were associated with decreased odds of influenza vaccine receipt (@gaskinEffectIncomeDisparities2023; @vashist_etal2022), and having health insurance coverage has also been shown to be associated with increased influenza vaccine coverage (@cambou_etal2021; @vashist_etal2022).
+Influenza vaccination rates were also found to be lower in rural than in urban areas among adults of all age groups and both genders (@jain_etal2022).
+However, there are still many variables of interest from the NHIS dataset that have yet to be evaluated in detail.
+Additionally, the periods of study from all prior research all end prior to the onset of the COVID-19 pandemic, which could have altered or heightened disparities even more.
 
@@
 
-The data source used for this project is the National Health Interview
-Survey (NHIS). It is a harmonized set of data beginning in 1963 with
-information on general health status, acute and chronic illness,
-functional limitations, access to care, insurance coverage, and health
-behaviors for the U.S. population. On average, the survey covers 100,000
-persons in 45,000 households each year. It is conducted in-person and by
-telephone, and its sample is designed to be representative of the
-civilian United States population (NHIS, About).
+The data source used for this project is the National Health Interview Survey (NHIS).
+It is a harmonized set of data beginning in 1963 with information on general health status, acute and chronic illness, functional limitations, access to care, insurance coverage, and health behaviors for the U.S.
+population.
+On average, the survey covers 100,000 persons in 45,000 households each year.
+It is conducted in-person and by telephone, and its sample is designed to be representative of the civilian United States population (@IPUMSNHIS.)
 
@@
 
-Individuals were eligible for the study if they were interviewed as part
-of the NHIS household survey between the years of 2012 and 2021. Table 1
-below provides a summary of demographic characteristics of our study
-population.
+Individuals were eligible for the study if they were interviewed as part of the NHIS household survey between the years of 2012 and 2021.
+Table 1 below provides a summary of demographic characteristics of our study population.
 
@@
 The raw data for the period of interest directly extracted from the NHIS
-IPUMS survey initially had had 792,916 observations. Researchers then
+IPUMS survey initially had 792,916 observations. NHIS Researchers then
 selected their predictors of interest to model the association between
@@
 
+```{r echo=FALSE, out.width = "50%", fig.align = "center"}
+knitr::include_graphics(here("results", "age & flu uptake distribution.png"))
+```
+
 \*\* Table 1: Demographic characteristics of the study population \*\*
@@
 
-```{r, warning=FALSE, message=FALSE}
+```{r, warning=FALSE, message=FALSE, echo=FALSE}
 result_table2 <- readRDS(here("results", "demographics_model.rds"))
@@
 
-```{r, warning=FALSE}
+```{r, warning=FALSE, echo=FALSE}
 result_table3 <- readRDS(here("results", "full_model.rds"))
@@
 
-\*\*
+### Model Evaluation
+
+#### ROC curve from the full model
+
+```{r echo=FALSE, out.width = "50%", fig.align = "center"}
+knitr::include_graphics(here("results", "AUC_full_model.png"))
+```
+
+#### ROC curve from the reduced model
+
+```{r echo=FALSE, out.width = "50%",  fig.align = "center"}
+knitr::include_graphics(here("results", "AUC_reduced_model.png"))
+```
 

### 2023-04-21 products/manuscript/Manuscript.qmd
Results, discussion, and conclusion updates

@@
 
-```{r, echo=FALSE, message=FALSE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 # load a few R packages
@@
 
-```{r, warning=FALSE, message=FALSE}
+```{r, warning=FALSE, message=FALSE, echo=FALSE}
 result_table1 <- readRDS(here("results", "table1_demographics.rds"))
@@
 
+The results of our analysis showed that age was significantly associated with flu vaccine uptake (OR = 1.031, 95% CI = 1.030-1.032, p < 0.001), with older individuals being more likely to have received the vaccine.
+The sex of the individual was also significantly associated with flu vaccine uptake (OR = 1.416, 95% CI = 1.384-1.448, p < 0.001), with females being more likely to have received the vaccine compared to males.
+Race was also significantly associated with flu vaccine uptake.
+Specifically, individuals who identified as black (OR = 0.790, 95% CI = 0.776-0.805, p < 0.001) were less likely to have received the vaccine compared to those who identified as white.
+Native Americans (OR = 1.284, 95% CI = 1.149-1.436, p = 0.0001) and Asians (OR = 1.141, 95% CI = 1.089-1.195, p < 0.001) were more likely to have received the vaccine than whites.
+Hispanic individuals (OR = 0.825, 95% CI = 0.806-0.845, p < 0.001) were less likely to have received the vaccine compared to non-Hispanic individuals.
+Education level was also significantly associated with flu vaccine uptake, with individuals who had higher levels of education being more likely to have received the vaccine.
+Specifically, individuals with some college education (OR = 1.172, 95% CI = 1.126-1.220, p < 0.001), a bachelor's degree (OR = 1.460, 95% CI = 1.402-1.520, p < 0.001), or a graduate degree (OR = 1.911, 95% CI = 1.818-2.007, p < 0.001) were more likely to have received the vaccine compared to those with no education.
+Individuals with family incomes of $50,000-$99,999 (OR = 1.140, 95% CI = 1.112-1.168, p < 0.001) and those with family incomes of $100,000 or more (OR = 1.417, 95% CI = 1.380-1.455, p < 0.001) were more likely to have received the vaccine compared to those with family incomes less than $50,000.
+In terms of health-related variables, individuals with asthma (OR = 1.398, 95% CI = 1.361-1.435, p < 0.001), cancer (OR = 1.322, 95% CI = 1.281-1.364, p < 0.001), and coronary heart disease (OR = 1.637, 95% CI = 1.560-1.717, p < 0.001) were more likely to have received the vaccine compared to those without these conditions.
+Finally, smoking status was significantly associated with flu vaccine uptake, with former smokers (OR = 1.059, 95% CI = 1.032-1.087, p = 0.0006) being more likely to receive the vaccine compared to nonsmoker, and current smokers (OR = 0.669, 95% CI = 0.643 - 0.696) being less likely to receive vaccine.
+Finally, former drinkers (OR = 1.059, 95% CI = 1.010-1.111, p = 0.0172) were more likely to receive the vaccine compared to nondrinkers, and this was statistically significant.
+However, current drinkers (OR = 0.978, 95% CI = 0.941 - 1.016) were less likely to receive vaccine compared to nondrinkers.
+Although this association was not statistically significant.
+
 ### Model Evaluation
@@
 
-*Summarize what you did, what you found and what it means.*
-
-[\*\*To be completed (PART 5)]{.underline}
+The results of our analysis indicate that several demographic and health-related factors are significantly associated with flu vaccine uptake.
+Our findings suggest that age, sex, race, education level, family income, and health status are all important predictors of vaccine uptake.
+Specifically, older individuals, females, individuals with higher education levels, and those with higher family incomes were more likely to have received the vaccine.
+On the other hand, individuals who identified as black or Hispanic were less likely to have received the vaccine, while Native Americans and Asians were more likely to have received it.
+In terms of health-related variables, individuals with asthma, cancer, and coronary heart disease were more likely to have received the vaccine compared to those without these conditions.
+Finally, our findings also suggest that smoking and drinking status are associated with vaccine uptake, with former smokers and drinkers more likely to have received the vaccine compared to nonsmokers and nondrinkers, respectively.
+These results have important implications for public health efforts aimed at increasing vaccine uptake, particularly among underserved and marginalized populations.
 
@@
 
-*Discuss what you perceive as strengths and limitations of your
-analysis.*
+While caution should be exercised when interpreting the significance of associations in observational studies due to potential confounding and reverse causality, this analysis provides valuable insights into identifying groups at risk for low influenza vaccine uptake over time.
+Limited studies have explored the relationship between a comprehensive list of factors and influenza vaccine coverage, and the large study population in this research improves statistical power, generalizability, precision, and accuracy of the findings.
+The design of the NHIS survey also allows for representative sampling of households, reducing sampling bias.
+Additionally, the survey questions were tailored to the month of each interview, enabling more detailed assessment of the seasonality of influenza vaccine schedule in the United States.
 
-[\*\*To be completed (PART 5)]{.underline}
+In terms of limitations, as noted above, observational studies cannot demonstrate causality.
+The reason for not receiving an influenza vaccine is not an NHIS variable, which could have provided more insight.
+It is also important to note that the primary outcome of vaccination within the past 12 months was self-reported, and there was no way to verify using medical records, meaning that there is a risk of reporting bias due to recall, social desirability, or other factors.
+The reliability of self-report compared to health records for influenza vaccine is not known.
+Lastly, It is also unclear to what extent the COVID-19 pandemic and antivaccine movements have impacted vaccine behavior and hesitancy, which could be explored in future qualitative studies.
 
@@
 
-*What are the main take-home messages?*
-
-[\*\*To be completed (PART 5)]{.underline}
+In conclusion, this study highlights the various demographic and health-related factors that are significantly associated with flu vaccine uptake.
+Understanding these predictors can help inform public health efforts aimed at increasing vaccine uptake, especially among marginalized populations.
+The findings suggest that targeted interventions may be necessary to address disparities in vaccine uptake among different racial and ethnic groups, as well as those with lower education levels and family incomes.
+Additionally, our findings on the association between smoking and drinking status and vaccine uptake may suggest the need for tailored messaging to encourage vaccination among these populations.
+Overall, these results can help guide public health interventions to promote vaccine uptake and ultimately reduce the burden of flu-related illness and hospitalizations.
 
 # References
-
-*Include citations in your Rmd file using bibtex, the list of references
-will automatically be placed at the end*
-
-[\*\*To be completed (PART 5)]{.underline}

### 2023-04-23 code/analysis_code/exploratory_analysis.qmd
More updates

@@
 ```{r, warning=FALSE, message=FALSE}
-#load needed packages. make sure they are installed.
 library(here) #for data loading/saving
@@
 library(haven)
-library("kableExtra") # For producing a nice table 1
+library(gridExtra) # To combine plots
+library("gtsummary") # For producing table 1
+#library("kableExtra") # For producing a nice table 1
 ```
@@
 
-### Print data
-
-```{r}
-print(mydata2)
-```
-
 # Data exploration through tables
@@
 ```
-### Checking names of data
+
+# Checking names of data
 ```{r}
@@
 
-## Data exploration through figures
+# Data exploration through figures
 
-###Box plots for the flu vaccine uptake and age
+```{r}
+getwd()
+```
+
+## Box plots for the flu vaccine uptake and age
 
@@
 p1 <- ggplot(mydata2, aes(x=flu_vaccine, y=age)) + 
-  geom_boxplot(fill = "gray") +
-  labs(title = "Distribution of Age per flu vaccine uptake", x = "Uptake of Flu Vaccine in the past 12 months", y = "Age") +  theme_classic()
+  geom_boxplot(fill = "#A8DADC") + scale_x_discrete(labels = c("No", "Yes")) +
+  labs(title = "Figure 1: Distribution of flu vaccine uptake by age", x = "Uptake of Flu Vaccine in the past 12 months", y = "Age") +  theme_classic() +
+  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
+        axis.title.x = element_text(size = 12, margin = margin(t = 12)),
+        axis.title.y = element_text(size = 12, margin = margin(r = 12)),
+        axis.text = element_text(size = 12),
+        legend.position = "none",
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank())
 
@@
 
+### Saving the plot in the result folder
 ```{r}
-figure_file = here("results","age & flu uptake distribution.png")
+figure_file = here("results","Figure 1: Age & flu uptake distribution.png")
 ggsave(filename = figure_file, plot=p1) 
@@
 
-### Barplot for flu vaccine uptake: 1 for no and 2 for yes
+### Uptake of flu vaccine by race
+```{r}
+p2 <- ggplot(mydata2, aes(x = flu_vaccine, fill = race)) +
+  geom_bar(position = "dodge", alpha = 0.8, color = "black") +
+  labs(title = "Figure 2: Distribution of flu vaccine uptake by race",
+       x = "Flu Vaccine Uptake in the Past 12 Months",
+       y = "Count") +
+  scale_x_discrete(labels = c("Did not take flu vaccine", "Did take flu vaccine")) +
+  scale_fill_manual(values = c("#A8DADC", "#F4A261", "#E76F51", "#E76F51"), labels = c("White", "Black", "Native American", "Asian")) +
+  theme_minimal() +
+  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
+        axis.title.x = element_text(size = 12, margin = margin(t = 12)),
+        axis.title.y = element_text(size = 12, margin = margin(r = 12)),
+        axis.text = element_text(size = 12),
+        legend.position = "top",
+        legend.title = element_blank(),
+        legend.text = element_text(size = 12),
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank())
 
+p2
+```
+
+### Saving the plot in the result folder
 ```{r}
-ggplot(mydata2, aes(flu_vaccine)) + 
-  geom_bar(width=0.7, fill="steelblue") + scale_x_discrete(labels=c('Did not take flu vaccine', 'Did take flu vaccine')) + theme_classic()
+figure_file = here("results","race & flu uptake distribution.png")
+ggsave(filename = figure_file, plot=p2) 
 ```
@@
 ```{r}
-ggplot(mydata2, aes(x= flu_vaccine, fill = sex)) + 
-  geom_bar(position = "dodge") + scale_x_discrete(labels=c('Did not take flu vaccine', 'Did take flu vaccine')) + scale_fill_discrete(labels=c('Male', 'Female')) +  theme_classic()
+p3 <- ggplot(mydata2, aes(x = flu_vaccine, fill = sex)) +
+  geom_bar(position = "dodge", alpha = 0.8, color = "black") +
+  labs(title = "Flu Vaccine Uptake by Sex",
+       x = "Flu Vaccine Uptake in the Past 12 Months",
+       y = "Count") +
+  scale_x_discrete(labels = c("Did not take flu vaccine", "Did take flu vaccine")) +
+  scale_fill_manual(values = c("#A8DADC", "#F4A261"), labels = c("Male", "Female")) +
+  theme_minimal() +
+  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
+        axis.title.x = element_text(size = 12, margin = margin(t = 12)),
+        axis.title.y = element_text(size = 12, margin = margin(r = 12)),
+        axis.text = element_text(size = 12),
+        legend.position = "top",
+        legend.title = element_blank(),
+        legend.text = element_text(size = 12),
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank())
 
+p3
+```
+
+### Saving the plot in the result folder
+```{r}
+figure_file = here("results","sex & flu uptake distribution.png")
+ggsave(filename = figure_file, plot=p3) 
 ```
@@
 ### Flu vaccine by smoking: 1 for nonsmoker, 2 for former smoker, 3 for current smoker
+```{r}
+p4 <- ggplot(mydata2, aes(x = flu_vaccine, fill = smoking_status)) +
+  geom_bar(position = "dodge", alpha = 0.8, color = "black") +
+  labs(title = "Flu Vaccine Uptake by Smoking Status",
+       x = "Flu Vaccine Uptake in the Past 12 Months",
+       y = "Count") +
+  scale_x_discrete(labels = c("Did not take flu vaccine", "Did take flu vaccine")) +
+  scale_fill_manual(values = c("#A8DADC", "#F4A261", "#E76F51"), labels = c("Nonsmoker", "Former Smoker", "Current Smoker")) +
+  theme_minimal() +
+  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
+        axis.title.x = element_text(size = 12, margin = margin(t = 12)),
+        axis.title.y = element_text(size = 12, margin = margin(r = 12)),
+        axis.text = element_text(size = 12),
+        legend.position = "top",
+        legend.title = element_blank(),
+        legend.text = element_text(size = 12),
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank())
 
-```{r, message=FALSE, warning=FALSE}
-ggplot(mydata2, aes(x = flu_vaccine, fill = smoking_status)) +
-  geom_histogram(stat="count") + scale_x_discrete(labels=c('Did not take flu vaccine', 'Did take flu vaccine')) + scale_fill_discrete(labels=c('Nonsmoker', 'Former Smoker', 'Current smoker')) +
-  ggtitle("Distribution of flu vaccine uptake and smoking status") + theme_classic()
+p4
+```
+
+### Saving the plot in the result folder
+```{r}
+figure_file = here("results","smoking status & flu uptake distribution.png")
+ggsave(filename = figure_file, plot=p4) 
 ```
@@
 
-```{r, message=FALSE, warning=FALSE}
-ggplot(mydata2, aes(x = flu_vaccine, fill = health_ins_status)) +
-  geom_histogram(stat="count") + scale_x_discrete(labels=c('Did not take flu vaccine', 'Did take flu vaccine')) + scale_fill_discrete(labels=c('Does not have health insurance', 'Does have health insurance')) +
-  ggtitle("Distribution of flu vaccine uptake by health insurance status") + theme_classic()
+```{r}
+p5 <- ggplot(mydata2, aes(x = flu_vaccine, fill = health_ins_status)) +
+  geom_bar(position = "dodge", alpha = 0.8, color = "black") +
+  labs(title = "Flu Vaccine Uptake by Insurance Status",
+       x = "Flu Vaccine Uptake in the Past 12 Months",
+       y = "Count") +
+  scale_x_discrete(labels = c("Did not take flu vaccine", "Did take flu vaccine")) +
+  scale_fill_manual(values = c("#A8DADC", "#F4A261"), labels = c('Does not have health insurance', 'Does have health insurance')) +
+  theme_minimal() +
+  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
+        axis.title.x = element_text(size = 12, margin = margin(t = 12)),
+        axis.title.y = element_text(size = 12, margin = margin(r = 12)),
+        axis.text = element_text(size = 12),
+        legend.position = "top",
+        legend.title = element_blank(),
+        legend.text = element_text(size = 12),
+        panel.grid.major = element_blank(),
+        panel.grid.minor = element_blank())
+
+p5
+```
+
+### Saving the plot in the result folder
+```{r}
+figure_file = here("results","Insurance status & flu uptake distribution.png")
+ggsave(filename = figure_file, plot=p5) 
 ```
@@
 
+```{r}
+tbl_summary(mydata2)
+```
+
+```{r}
+a <- mydata2 %>% select(age, sex)
+b <- tbl_summary(a)
+```
+
 ## Full table for demogrophics characteristics
@@
 library(table1)
-table_dems <- (table1(~ factor(sex) + age + factor(race) + factor(hispanic) + factor(education) + factor(fam_income) + factor(employ_status) + factor(paid_sick) + factor(health_ins_status) + factor(med_access) + factor(med_care_delayed_cost) + factor(health_status) + factor(bmi) + factor(asthma) + factor(cancer) + factor(chd) + factor(alcohol_status) + factor(smoking_status)| flu_vaccine, data=mydata2))
+table_dems <- (table1(~ sex + age + race + hispanic + education + fam_income + employ_status + paid_sick + health_ins_status + med_access + med_care_delayed_cost + health_status + bmi + asthma + cancer + chd + alcohol_status + smoking_status| flu_vaccine, data=mydata2))
 
@@
 
+```{r}
+install.packages("kableExtra", dependencies = TRUE)
+```
+
 ### Saving table 1 in the results section

### 2023-04-23 products/manuscript/Manuscript.qmd
More updates

@@
 
-### Description of data and data source
+## Description of data and data source
 
@@
 
-#### Study population
+## Study population
 
@@
 
-### Statistical analysis
+## Statistical analysis
 
@@
 
-### Exploratory/Descriptive analysis
+## Exploratory/Descriptive analysis
 
-```{r echo=FALSE, out.width = "50%", fig.align = "center"}
+```{r echo=FALSE, out.width = "10%",  fig.align = "center"}
 knitr::include_graphics(here("results", "age & flu uptake distribution.png"))
+```
+
+```{r echo=FALSE, out.width = "10%",  fig.align = "center"}
+knitr::include_graphics(here("results", "race & flu uptake distribution.png"))
 ```
@@
 
-### Loading Table 1 from the Exploratory Analysis
+## Loading Table 1 from the Exploratory Analysis
 
@@
 
-### Full analysis
+## Full analysis
 
@@
 
-### Model 2: Using all our predictors of interest in the model with flu vaccine uptake as our outcome
+## Model 2: Using all our predictors of interest in the model with flu vaccine uptake as our outcome
 
@@
 
-### Model Evaluation
+## Model Evaluation
 
-#### ROC curve from the full model
+### ROC curve from the full model
 
@@
 
-#### ROC curve from the reduced model
+### ROC curve from the reduced model
 
@@
 
-### Summary and Interpretation
+## Summary and Interpretation
 
@@
 
-### Strengths and Limitations
+## Strengths and Limitations
 
@@
 
-### Conclusions
+# Conclusions
 

### 2023-04-27 code/analysis_code/exploratory_analysis.qmd
add review from Weifan Wu

@@
 # Setup
-
 ```{r, warning=FALSE, message=FALSE}
@@
 library(gridExtra) # To combine plots
-library("gtsummary") # For producing table 1
+library(gtsummary) # For producing table 1
 #library("kableExtra") # For producing a nice table 1

### 2023-04-27 review_Weifan_Wu.qmd
add review from Weifan Wu

---
title: Project Review
author: YOUR NAME
date: "`r file.mtime(knitr::current_input())`"
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: Which groups are at highest risk for low uptake of flu vaccine?

Name of project author(s): Chris Okitondo and Irene Cavros

Name of project reviewer: Weifan Wu

## Background, Context and Motivation
How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

The context of this project is very detailed. The introduction covers the clinical presentation of influenza and the importance of vaccine coverage. It goes on to discuss the previous studies conducted to examine the potential contribution factors to low-coverage of the vaccination. Therefore, the background information is thorough and the motivation of the study is clear. However, at the end of the introduction, it would be better to briefly talk about the objective of this study to further clarify your research goals and set expectations for the reader.

### Summary assessment (PICK ONE, DELETE THE OTHERS)
* strong contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

The question description is missing.

### Summary assessment
* question/hypotheses unclear

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

The data description is clear and the source is provided. The codebook is available in the readme.md file placed in the raw_data folder. The in-text citation for the data source, however, is incorrect.

### Summary assessment

* source and overall structure of data well explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

The data processing step is well-explained. Each name of the variables and the levels under each variable were renamed and reorganized into an intuitive way so that the projector owners and audiences (including myself) can easily follow along.
The EDA part is fairly well-done. However, there is still room for enhancement. In Figure 2, for better presentation, it would be more appropriate to use the percentage of vaccinated individuals within each racial group on the y-axis, rather than raw counts, due to the significant differences in total population sizes among the races. Moreover, the majority of the presented graphs and tables focus on univariate relationships, which may overlook crucial confounding factors that could influence the results.

### Summary assessment
* essentially no weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

The analysis part is well-explained and presented in a good layout. However,the project only employs a logistic regression model for analysis. It would be beneficial to incorporate one or two machine learning approaches that we learnt from this course to construct the predictive model. By doing so, the performance of each model can be compared, allowing for the selection of the best model to generate the final results.

### Summary assessment

* defensible but not optimal analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments
The plots and tables for EDA are really nice. However, in the full analysis part, organizing the results into tables using the kable() function from the knitr package or the gt() function from the gt package would enhance their presentations. Also the results could be written in fully reproducible manner using inline R code like OR= r tabel1$estimate[2] instead of manually typing the numbers. This practice would ensure that any updates or changes to the data are automatically reflected in the reported findings, reducing the risk of errors and inconsistencies.

### Summary assessment
* results are presented ok, with room for improvement

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments
The discussion part is well-presented and consistent with the results. Confounding factors are addressed in the limitations portion of the report. However, incorporating comparisons with additional studies could further strengthen and explain the project's results.

### Summary assessment
* strong, complete and clear discussion

## Further comments

The package loading section of manuscript can be excluded in the final word document by using `include=FALSE` within the r code chunk.

# Overall project content evaluation
Evaluate overall features of the project  by filling in the sections below.

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments
The project is well structured and labelled with appropriate names. There is no irrelevant or redundant file. It was fairly easy to reproduce everything by following the instructions.

### Summary assessment
* well structured

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

Each step of the whole analysis is well-explained. There are plenty of comments justifying and stating the purpose of each line of code.

### Summary assessment

* fully and well documented

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

All results are fully reproducible and the readme.me file under each folder provides clear guidance.

### Summary assessment
* fully reproducible without issues

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

As previously noted, the data analysis exclusively employed a logistic regression model, and adding alternative models could lead to more comprehensive analysis. Additionally, the study did not present any research questions or hypotheses before conducting the investigation, which would have been beneficial for guiding the analysis.

### Summary assessment
* decent level of thoroughness

### 2023-05-05 code/analysis_code/exploratory_analysis.qmd
More edits v2

@@
 library(gtsummary) # For producing table 1
-#library("kableExtra") # For producing a nice table 1
+library("kableExtra") # For producing a nice table 1
+library(table1)
 ```
@@
 names(mydata2)
-```
-
-# Data exploration through figures
-
-```{r}
-getwd()
 ```
@@
 ```{r, warning=FALSE, message=FALSE}
-library(table1)
 (table1(~ factor(sex) + age + factor(race) + factor(hispanic) | flu_vaccine, data=mydata2))
@@
 ```{r, warning=FALSE, message=FALSE}
-library(table1)
 (table1(~ factor(education) + factor(fam_income) | flu_vaccine, data=mydata2))
@@
 ```{r, warning=FALSE, message=FALSE}
-library(table1)
 (table1(~ factor(employ_status) + factor(paid_sick) | flu_vaccine, data=mydata2))
@@
 ```{r, warning=FALSE, message=FALSE}
-library(table1)
 (table1(~ factor(health_ins_status) + factor(med_access) + factor(med_care_delayed_cost) | flu_vaccine, data=mydata2))
@@
 ```{r, warning=FALSE, message=FALSE}
-library(table1)
 (table1(~ factor(health_status) + factor(bmi) + factor(asthma) + factor(cancer) + factor(chd) + factor(alcohol_status) + factor(smoking_status) | flu_vaccine, data=mydata2))
@@
 ```{r, warning=FALSE, message=FALSE}
-library(table1)
 (table1(~factor(worry_freq) + factor(dep_freq) | flu_vaccine, data=mydata2))
@@
 
-```{r}
-tbl_summary(mydata2)
+## Full table for demogrophics characteristics
+```{r, warning=FALSE, message=FALSE}
+table_dems <- (table1(~ sex + age + race + hispanic + education + fam_income + employ_status + paid_sick + health_ins_status + med_access + med_care_delayed_cost + health_status + bmi + asthma + cancer + alcohol_status + smoking_status| flu_vaccine, data=mydata2))
+
+table_dems
+
 ```
@@
 ```{r}
-a <- mydata2 %>% select(age, sex)
-b <- tbl_summary(a)
-```
 
-## Full table for demogrophics characteristics
-```{r, warning=FALSE, message=FALSE}
 library(table1)
-table_dems <- (table1(~ sex + age + race + hispanic + education + fam_income + employ_status + paid_sick + health_ins_status + med_access + med_care_delayed_cost + health_status + bmi + asthma + cancer + chd + alcohol_status + smoking_status| flu_vaccine, data=mydata2))
 
-table_dems
+table_dems6 <- table1(~ sex + age + race + hispanic + education + fam_income + employ_status + paid_sick + health_ins_status + med_access + med_care_delayed_cost + health_status + bmi + asthma + cancer + alcohol_status + smoking_status | flu_vaccine, data = mydata2,
+                     label = list(
+                       sex = "Sex",
+                       age = "Age",
+                       race = "Race",
+                       hispanic = "Hispanic Ethnicity",
+                       education = "Education",
+                       fam_income = "Family Income",
+                       employ_status = "Employment Status",
+                       paid_sick = "Paid Sick Leave",
+                       health_ins_status = "Health Insurance Status",
+                       med_access = "Medical Access",
+                       med_care_delayed_cost = "Delayed Medical Care Cost",
+                       health_status = "Health Status",
+                       bmi = "Body Mass Index",
+                       asthma = "Asthma",
+                       cancer = "Cancer",
+                       alcohol_status = "Alcohol Consumption",
+                       smoking_status = "Smoking Status",
+                       flu_vaccine = "Flu Vaccine"
+                     ),
+                     show_na = TRUE
+                    )
 
+table_dems6
 ```
@@
 
-```{r}
-install.packages("kableExtra", dependencies = TRUE)
-```
-
 ### Saving table 1 in the results section
@@
 ```
+
+######## CLEANING TABLE 1 #######
+
+```{r}
+ (table1(~ sex + age + race + hispanic + education + fam_income + employ_status + paid_sick + health_ins_status + med_access + med_care_delayed_cost + health_status + bmi + asthma + cancer + chd + alcohol_status + smoking_status| flu_vaccine, data=mydata2))
+
+```
+
+```{r}
+mydata2 %>% tbl_summary(include = c(age, sex))
+```
+
+```{r}
+ tbl_summary(
+    mydata2,
+    include = c(sex, age, race, hispanic, education, fam_income, employ_status,    paid_sick, health_ins_status, med_access, med_care_delayed_cost, health_status, bmi, asthma, cancer, alcohol_status, smoking_status),
+    by = flu_vaccine, # split table by group
+    missing = "no" # don't list missing data separately
+  ) %>%
+  add_n() %>% # add column with total number of non-missing observations
+  add_p() %>% # test for a difference between groups
+  modify_header(label = "**Variable**") %>% # update the column header
+  bold_labels()
+```
+
+### Cleaner version of Table 1: Demographic Characteristics
+```{r}
+table_dems_v3 <- mydata2 %>%
+  tbl_summary(
+    statistic = list(all_categorical() ~ "{n} ({p}%)", all_continuous() ~ "{mean} ({sd})"),
+    by = flu_vaccine,
+    missing = "no",
+    label = list(
+      sex ~ "Sex",
+      age ~ "Age",
+      race ~ "Race",
+      hispanic ~ "Hispanic Ethnicity",
+      education ~ "Education",
+      fam_income ~ "Family Income",
+      employ_status ~ "Employment Status",
+      paid_sick ~ "Paid Sick Leave",
+      health_ins_status ~ "Health Insurance Status",
+      med_access ~ "Medical Access",
+      med_care_delayed_cost ~ "Delayed Medical Care Cost",
+      health_status ~ "Health Status",
+      bmi ~ "Body Mass Index",
+      asthma ~ "Asthma",
+      cancer ~ "Cancer",
+      alcohol_status ~ "Alcohol Consumption",
+      smoking_status ~ "Smoking Status"
+    )
+  ) %>%
+  bold_labels() %>%
+  add_overall() %>%
+  add_n() %>%
+  add_p() %>%
+  modify_header(label = "**Variable**")
+
+table_dems_v3
+```

### 2023-05-05 code/analysis_code/statistical_analysis.qmd
More edits v2

---
title: "Model Fitting"
author: "Chris Okitondo and Irene Cavros"
date: "03/13/2023"
output: html_document
---

In this section, we will undergo the following two tasks:

-   Fit various bivariate logistic models to the categorical outcome (taking the flu vaccine) with individual important predictors of interest.

-   Fit logistic models to the categorical outcome (taking the flu vaccine) using all (important) predictors of interest.

### Setup

```{r, warning=FALSE, message=FALSE}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
library(tidyverse)
library(tidymodels)
library(performance)
```

### Load data

```{r}
#Path to data. 
data_location <- here::here("data","processed_data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

### View data
```{r}
View(mydata)
```

### Take another quick look at the data

```{r}
glimpse(mydata)
```

### Ensure we are ignoring the NA values in the dataset
```{r}
mydata2 <- na.omit(mydata)
```

### Check variable names

```{r}
names(mydata2)
```

### Convert scientific notation into decimals

*This will help us more easily determine which variables are significant.*

```{r}
options(digits=9)
```

### Set up logistic regression

```{r}
log_mod <- logistic_reg() %>%
  set_engine("glm")
```

# Fitting bivariate logistic models

### Using flu uptake as the outcome and age as predictor

```{r}
log_fit_age <- log_mod %>% fit(as.factor(flu_vaccine) ~ age, data = mydata2)
tidy(log_fit_age)
```

### Using flu uptake as the outcome and sex as predictor

```{r}
log_fit_sex <- log_mod %>% fit(as.factor(flu_vaccine) ~ sex, data = mydata2)
tidy(log_fit_sex)
```

### Using flu uptake as the outcome and race as predictor

```{r}
log_fit_race <- log_mod %>% fit(as.factor(flu_vaccine) ~ race, data = mydata2)
tidy(log_fit_race)
```

*This shows that the black population and Native Americans are less likely to take the flu vaccine compared to white people. However, asians are more likely to take the flu vaccine compared to whites.*

### Using flu uptake as the outcome and education as predictor

```{r}
log_fit_education <- log_mod %>% fit(as.factor(flu_vaccine) ~ education, data = mydata2)
tidy(log_fit_education)
```

*This results show that the higher the degree someone has, the more likely they are to take the flu vaccine.*

### Using flu uptake as outcome and family income as predictor

```{r}
log_fit_fam_income <- log_mod %>% fit(as.factor(flu_vaccine) ~ fam_income, data = mydata2)
tidy(log_fit_fam_income)
```

*The results show that the higher their income, the greater the likelihood for someone to take the flu vaccine.*

### Using flu uptake as outcome and health insurance status as predictor

```{r}
log_fit_health_ins <- log_mod %>% fit(as.factor(flu_vaccine) ~ health_ins_status, data = mydata2)
tidy(log_fit_health_ins)
```

*Those with insurance coverage are more likely to take the flu vaccine.*

### Using flu uptake as outcome and medical care access as predictor

```{r}
log_fit_medical_care <- log_mod %>% fit(as.factor(flu_vaccine) ~ med_access, data = mydata2)
tidy(log_fit_medical_care)
```

*If they have at least one usual place to seek medical care, they are more likely to take the flu vaccine.*

```{r}
names(mydata2)
```

# Fitting multivariate logistic models

### Categorical outcome as flu vaccine with some demographic variables as predictors of interest

```{r}
log_fit_dems <- log_mod %>% fit(as.factor(flu_vaccine) ~ age + sex + race + hispanic + education + fam_income, data = mydata2)

dems_model <- tidy(log_fit_dems)
dems_model

saveRDS(dems_model, file = (here("results", "demographics_model.rds")))
```

### Categorical outcome as flu vaccine with asthma, cancer, and CHD as predictors of interest

```{r}
log_fit_medical_cond <- log_mod %>% fit(as.factor(flu_vaccine) ~  asthma+ chd + cancer, data = mydata2)

tidy(log_fit_medical_cond)
```

### Categorical outcome as flu vaccine with smoking and alcohol status as predictors of interest

```{r}
log_fit_health_behaviors <- log_mod %>% fit(as.factor(flu_vaccine) ~  smoking_status+ alcohol_status, data = mydata2)

tidy(log_fit_health_behaviors)
```

*This results show that former smokers and former drinkers are more likely to get the flu vaccine compared to the population who never smoked or drank. The active smokers are less likely to take the vaccine compared to those who do not drink. There was no statistical difference between the active drinkers and those who do not drink in terms of their vaccine uptake.*

### Categorical outcome as flu vaccine with smoking and alcohol as predictors of interest

```{r}
log_fit_full <- log_mod %>% fit(as.factor(flu_vaccine) ~  age + sex + race + hispanic + education + fam_income + asthma + cancer + smoking_status+ alcohol_status, data = mydata2)

table_full <- tidy(log_fit_full)
table_full

saveRDS(table_full, file = (here("results", "full_model.rds")))
```

*The results show that those who graduated from high school did not statistically differ from those who never graduated for their flu vaccine uptake. Also, current drinkers did not statically differ from those who never drank. All other predictors were found statistically significant.*

```{r}
log_fit_full2 <- log_mod %>% fit(as.factor(flu_vaccine) ~  age + sex + race + hispanic + education + fam_income + asthma + cancer + smoking_status+ alcohol_status, data = mydata2)

table_full <- tidy(log_fit_full2)
table_full

```

```{r}
log_fit_full <- log_mod %>% fit(as.factor(flu_vaccine) ~  age + sex + race + hispanic + education + fam_income + asthma + cancer +  smoking_status+ alcohol_status, data = mydata2)

t1 <- tbl_regression(log_fit_full, exponentiate = TRUE)
t1
```

```{r}

library(knitr)

log_fit_full <- log_mod %>% fit(as.factor(flu_vaccine) ~  age + sex + race + hispanic + education + fam_income + asthma + cancer + smoking_status + alcohol_status, data = mydata2)

t1 <- tbl_regression(log_fit_full, exponentiate = TRUE) %>%
  add_column(".rownames" = rownames(t1), .before = 1) %>%
  rename(
    "Variable" = term,
    "Odds Ratio" = estimate,
    "95% CI" = conf.low:conf.high,
    "p-value" = p.value
  ) %>%
  select(-c(".rownames", ".keep")) %>%
  mutate(
    Variable = case_when(
      Variable == "age" ~ "Age",
      Variable == "sex" ~ "Sex",
      Variable == "race" ~ "Race",
      Variable == "hispanic" ~ "Hispanic Ethnicity",
      Variable == "education" ~ "Education",
      Variable == "fam_income" ~ "Family Income",
      Variable == "asthma" ~ "Asthma",
      Variable == "cancer" ~ "Cancer",
      Variable == "smoking_status" ~ "Smoking Status",
      Variable == "alcohol_status" ~ "Alcohol Consumption",
      TRUE ~ Variable
    )
  ) %>%
  kable(format = "markdown", align = "c") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```

```{r}
library(knitr)
kable(table_full, 
      caption = "Multivariate Logistic Regression of Influenza Vaccination Status") %>%
  kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

### 2023-05-05 products/manuscript/Manuscript.qmd
More edits v2

@@
 
-Influenza, or flu, is an extremely contagious respiratory illness caused by a virus.
-It can cause a range of symptoms, including fever, cough, sore throat, body aches, headache, chills, and fatigue (@cdc2022c.) It can be particularly dangerous for certain groups, such as older adults, young children, and individuals with underlying health issues, often resulting in severe illness or even death.
-Although influenza vaccinations are widely accessible, many people in the United States do not receive them as recommended.
-This study examines influenza vaccination uptake among US adults from 2012-2021.
-Using a subset of the National Health Interview Survey (NHIS) data set, researchers provide important insights into which groups are at highest risk for under-coverage of flu vaccine in order to better inform targeted public health interventions for improving uptake.
+Influenza, or flu, is an extremely contagious respiratory illness caused
+by a virus. It can cause a range of symptoms, including fever, cough,
+sore throat, body aches, headache, chills, and fatigue (CDC (@2022c).)
+It can be particularly dangerous for certain groups, such as older
+adults, young children, and individuals with underlying health issues,
+often resulting in severe illness or even death. Although influenza
+vaccinations are widely accessible, many people in the United States do
+not receive them as recommended. This study examines influenza
+vaccination uptake among US adults from 2012-2021. Using a subset of the
+National Health Interview Survey (NHIS) data set, researchers provide
+important insights into which groups are at highest risk for
+under-coverage of flu vaccine in order to better inform targeted public
+health interventions for improving uptake.
 
@@
 
-Influenza, commonly known as the flu, is a contagious respiratory illness which comes from influenza viruses.
-It can cause mild to severe illness, and in some cases, it can lead to hospitalization or even death.
-Symptoms of the flu include fever, cough, sore throat, body
-aches, headache, chills, and fatigue (@cdc2022.) It can spread easily from person to person through droplets that are released infected individuals cough, sneeze, or talk.
+Influenza, commonly known as the flu, is a contagious respiratory
+illness which comes from influenza viruses. It can cause mild to severe
+illness, and in some cases, it can lead to hospitalization or even
+death. Symptoms of the flu include fever, cough, sore throat, body
+aches, headache, chills, and fatigue (@CDC (2022a).) It can spread
+easily from person to person through droplets that are released infected
+individuals cough, sneeze, or talk.
 
-Vaccines are a crucial tool in preventing the spread of infectious diseases, including the flu.
-On an individual level, the best way to protect against the flu is by getting vaccinated each year.
-At the population level, high vaccine coverage is a powerful strategy for reducing spread of the virus (@cdc2022c.)
+Vaccines are a crucial tool in preventing the spread of infectious
+diseases, including the flu. On an individual level, the best way to
+protect against the flu is by getting vaccinated each year. At the
+population level, high vaccine coverage is a powerful strategy for
+reducing spread of the virus (@CDC (2022c).)
 
-The U.S.
-Centers for Disease Control and Prevention (CDC) recommends that everyone six months of age and older receive a flu vaccine each year, with rare exceptions (@cdc2022c.) Influenza is a constantly evolving virus that mutates rapidly, and uncontrolled spread gives rise to many different strains, so new vaccines are developed each year to protect against the most common strains of the virus that are expected to circulate during the upcoming flu season.
-It is especially important for individuals who are at higher risk of severe illness or complications from the flu, such as young children, older adults, pregnant women, and individuals with underlying medical conditions.
+The U.S. Centers for Disease Control and Prevention (@CDC) recommends
+that everyone six months of age and older receive a flu vaccine each
+year, with rare exceptions (@CDC (2022c).) Influenza is a constantly
+evolving virus that mutates rapidly, and uncontrolled spread gives rise
+to many different strains, so new vaccines are developed each year to
+protect against the most common strains of the virus that are expected
+to circulate during the upcoming flu season. It is especially important
+for individuals who are at higher risk of severe illness or
+complications from the flu, such as young children, older adults,
+pregnant women, and individuals with underlying medical conditions.
 
-In terms of the availability and cost, there have been significant changes for flu vaccines over the past 50 years (@InfluenzaHistoricTimeline2022.) In the 1960s, flu vaccines were not widely available and were primarily given to people considered to be at high risk for complications.
-The cost of the vaccine was relatively high and was not covered by most insurance plans.
-In the 1970s and 1980s, flu vaccines became more widely available and were recommended for more people, including children and healthy adults.
-The cost of the vaccine began to decrease as more people received it, and insurance coverage for the vaccine became more common.
-In the 1990s and 2000s, the availability of flu vaccines continued to increase, and the cost of the vaccine became more affordable for a larger proportion of the U.S.
-population.
+In terms of the availability and cost, there have been significant
+changes for flu vaccines over the past 50 years (@Influenza Historic
+Timeline \| Pandemic Influenza (Flu) \| CDC (2022).) In the 1960s, flu
+vaccines were not widely available and were primarily given to people
+considered to be at high risk for complications. The cost of the vaccine
+was relatively high and was not covered by most insurance plans. In the
+1970s and 1980s, flu vaccines became more widely available and were
+recommended for more people, including children and healthy adults. The
+cost of the vaccine began to decrease as more people received it, and
+insurance coverage for the vaccine became more common. In the 1990s and
+2000s, the availability of flu vaccines continued to increase, and the
+cost of the vaccine became more affordable for a larger proportion of
+the U.S. population.
 
-Today, flu vaccines are frequently available in doctor's offices, pharmacies, community clinics, workplace clinics, school-based clinics, and local health departments, among others.
-They are often covered by insurance plans, which has increased their accessibility.
-Many public health clinics and community organizations also offer flu vaccines at low or no cost, in order to better target historically unreached populations.
+Today, flu vaccines are frequently available in doctor's offices,
+pharmacies, community clinics, workplace clinics, school-based clinics,
+and local health departments, among others. They are often covered by
+insurance plans, which has increased their accessibility. Many public
+health clinics and community organizations also offer flu vaccines at
+low or no cost, in order to better target historically unreached
+populations.
 
-Despite the availability and recommendation of influenza vaccines, many people in the United States still do not receive them as recommended.
-In fact, the CDC reported that only 49.2% of adults aged 18 years and older received a flu vaccine during the 2020-2021 flu season (@cdc2022b.) This low uptake of influenza vaccine highlights the need for understanding the factors that contribute to under-coverage of the vaccine in certain populations.
+Despite the availability and recommendation of influenza vaccines, many
+people in the United States still do not receive them as recommended. In
+fact, the CDC reported that only 49.2% of adults aged 18 years and older
+received a flu vaccine during the 2020-2021 flu season (@CDC (2022b).)
+This low uptake of influenza vaccine highlights the need for
+understanding the factors that contribute to under-coverage of the
+vaccine in certain populations.
 
-Racial/ethnic minorities have a long history of reduced vaccination rates relative to white Americans (@khan_etal2018.) Since 2010, flu vaccination coverage has been consistently lower among Black, Hispanic, and AI/AN adults.
-During the 2021--2022 season, flu vaccination coverage was 54% among White adults, 42% among Black adults, 38% among Hispanic adults, and 41% among AI/AN adults (@cdc2022b.) There are many reasons for these inequities, including lack of access to health care and insurance, missed opportunities to vaccinate, and misinformation and distrust.
+Racial and ethnic minorities have a long history of reduced vaccination
+rates relative to white Americans (@Khan, @Hall, @Tanner, & @Marlow
+(@2018).) Since 2010, flu vaccination coverage has been consistently
+lower among Black, Hispanic, and AI/AN adults. During the 2021--2022
+season, flu vaccination coverage was 54% among White adults, 42% among
+Black adults, 38% among Hispanic adults, and 41% among AI/AN adults
+(@CDC (2022b).) There are many reasons for these inequities, including
+lack of access to health care and insurance, missed opportunities to
+vaccinate, and misinformation and distrust.
 
-In prior studies, lower income levels were associated with decreased odds of influenza vaccine receipt (@gaskinEffectIncomeDisparities2023; @vashist_etal2022), and having health insurance coverage has also been shown to be associated with increased influenza vaccine coverage (@cambou_etal2021; @vashist_etal2022).
-Influenza vaccination rates were also found to be lower in rural than in urban areas among adults of all age groups and both genders (@jain_etal2022).
-However, there are still many variables of interest from the NHIS dataset that have yet to be evaluated in detail.
-Additionally, the periods of study from all prior research all end prior to the onset of the COVID-19 pandemic, which could have altered or heightened disparities even more.
+In prior studies, lower income levels were associated with decreased
+odds of influenza vaccine receipt (@Gaskin, Woods, Ghosh, Watson, &
+Huber (2023); Vashist, Choi, & Patel (2022)), and having health
+insurance coverage has also been shown to be associated with increased
+influenza vaccine coverage (Cambou, Copeland, Nielsen-Saines, & Macinko
+(2021); Vashist et al. (2022)). Influenza vaccination rates were also
+found to be lower in rural than in urban areas among adults of all age
+groups and both genders (Jain et al. (2022)). However, there are still
+many variables of interest from the NHIS dataset that have yet to be
+evaluated in detail. Additionally, the periods of study from all prior
+research all end prior to the onset of the COVID-19 pandemic, which
+could have altered or heightened disparities even more.
+
+## Question:
+
+Looking at whether study participants received any influenza vaccine in
+the past twelve month, this study aims to answer the following question:
+
+Which groups are at highest risk for low uptake of the influenza
+vaccine?
 
@@
 
-The data source used for this project is the National Health Interview Survey (NHIS).
-It is a harmonized set of data beginning in 1963 with information on general health status, acute and chronic illness, functional limitations, access to care, insurance coverage, and health behaviors for the U.S.
-population.
-On average, the survey covers 100,000 persons in 45,000 households each year.
-It is conducted in-person and by telephone, and its sample is designed to be representative of the civilian United States population (@IPUMSNHIS.)
+The data source used for this project is the National Health Interview
+Survey (NHIS). It is a harmonized set of data beginning in 1963 with
+information on general health status, acute and chronic illness,
+functional limitations, access to care, insurance coverage, and health
+behaviors for the U.S. population. On average, the survey covers 100,000
+persons in 45,000 households each year. It is conducted in-person and by
+telephone, and its sample is designed to be representative of the
+civilian United States population (@IPUMSNHIS.)
 
@@
 
-Individuals were eligible for the study if they were interviewed as part of the NHIS household survey between the years of 2012 and 2021.
-Table 1 below provides a summary of demographic characteristics of our study population.
+Individuals were eligible for the study if they were interviewed as part
+of the NHIS household survey between the years of 2012 and 2021. Table 1
+below provides a summary of demographic characteristics of our study
+population.
 
@@
 
-The results of our analysis showed that age was significantly associated with flu vaccine uptake (OR = 1.031, 95% CI = 1.030-1.032, p < 0.001), with older individuals being more likely to have received the vaccine.
-The sex of the individual was also significantly associated with flu vaccine uptake (OR = 1.416, 95% CI = 1.384-1.448, p < 0.001), with females being more likely to have received the vaccine compared to males.
-Race was also significantly associated with flu vaccine uptake.
-Specifically, individuals who identified as black (OR = 0.790, 95% CI = 0.776-0.805, p < 0.001) were less likely to have received the vaccine compared to those who identified as white.
-Native Americans (OR = 1.284, 95% CI = 1.149-1.436, p = 0.0001) and Asians (OR = 1.141, 95% CI = 1.089-1.195, p < 0.001) were more likely to have received the vaccine than whites.
-Hispanic individuals (OR = 0.825, 95% CI = 0.806-0.845, p < 0.001) were less likely to have received the vaccine compared to non-Hispanic individuals.
-Education level was also significantly associated with flu vaccine uptake, with individuals who had higher levels of education being more likely to have received the vaccine.
-Specifically, individuals with some college education (OR = 1.172, 95% CI = 1.126-1.220, p < 0.001), a bachelor's degree (OR = 1.460, 95% CI = 1.402-1.520, p < 0.001), or a graduate degree (OR = 1.911, 95% CI = 1.818-2.007, p < 0.001) were more likely to have received the vaccine compared to those with no education.
-Individuals with family incomes of $50,000-$99,999 (OR = 1.140, 95% CI = 1.112-1.168, p < 0.001) and those with family incomes of $100,000 or more (OR = 1.417, 95% CI = 1.380-1.455, p < 0.001) were more likely to have received the vaccine compared to those with family incomes less than $50,000.
-In terms of health-related variables, individuals with asthma (OR = 1.398, 95% CI = 1.361-1.435, p < 0.001), cancer (OR = 1.322, 95% CI = 1.281-1.364, p < 0.001), and coronary heart disease (OR = 1.637, 95% CI = 1.560-1.717, p < 0.001) were more likely to have received the vaccine compared to those without these conditions.
-Finally, smoking status was significantly associated with flu vaccine uptake, with former smokers (OR = 1.059, 95% CI = 1.032-1.087, p = 0.0006) being more likely to receive the vaccine compared to nonsmoker, and current smokers (OR = 0.669, 95% CI = 0.643 - 0.696) being less likely to receive vaccine.
-Finally, former drinkers (OR = 1.059, 95% CI = 1.010-1.111, p = 0.0172) were more likely to receive the vaccine compared to nondrinkers, and this was statistically significant.
-However, current drinkers (OR = 0.978, 95% CI = 0.941 - 1.016) were less likely to receive vaccine compared to nondrinkers.
-Although this association was not statistically significant.
+The results of our analysis showed that age was significantly associated
+with flu vaccine uptake (OR = 1.031, 95% CI = 1.030-1.032, p \< 0.001),
+with older individuals being more likely to have received the vaccine.
+The sex of the individual was also significantly associated with flu
+vaccine uptake (OR = 1.416, 95% CI = 1.384-1.448, p \< 0.001), with
+females being more likely to have received the vaccine compared to
+males. Race was also significantly associated with flu vaccine uptake.
+Specifically, individuals who identified as black (OR = 0.790, 95% CI =
+0.776-0.805, p \< 0.001) were less likely to have received the vaccine
+compared to those who identified as white. Native Americans (OR = 1.284,
+95% CI = 1.149-1.436, p = 0.0001) and Asians (OR = 1.141, 95% CI =
+1.089-1.195, p \< 0.001) were more likely to have received the vaccine
+than whites. Hispanic individuals (OR = 0.825, 95% CI = 0.806-0.845, p
+\< 0.001) were less likely to have received the vaccine compared to
+non-Hispanic individuals. Education level was also significantly
+associated with flu vaccine uptake, with individuals who had higher
+levels of education being more likely to have received the vaccine.
+Specifically, individuals with some college education (OR = 1.172, 95%
+CI = 1.126-1.220, p \< 0.001), a bachelor's degree (OR = 1.460, 95% CI =
+1.402-1.520, p \< 0.001), or a graduate degree (OR = 1.911, 95% CI =
+1.818-2.007, p \< 0.001) were more likely to have received the vaccine
+compared to those with no education. Individuals with family incomes of
+\$50,000-\$99,999 (OR = 1.140, 95% CI = 1.112-1.168, p \< 0.001) and
+those with family incomes of \$100,000 or more (OR = 1.417, 95% CI =
+1.380-1.455, p \< 0.001) were more likely to have received the vaccine
+compared to those with family incomes less than \$50,000. In terms of
+health-related variables, individuals with asthma (OR = 1.398, 95% CI =
+1.361-1.435, p \< 0.001), cancer (OR = 1.322, 95% CI = 1.281-1.364, p \<
+0.001), and coronary heart disease (OR = 1.637, 95% CI = 1.560-1.717, p
+\< 0.001) were more likely to have received the vaccine compared to
+those without these conditions. Finally, smoking status was
+significantly associated with flu vaccine uptake, with former smokers
+(OR = 1.059, 95% CI = 1.032-1.087, p = 0.0006) being more likely to
+receive the vaccine compared to nonsmoker, and current smokers (OR =
+0.669, 95% CI = 0.643 - 0.696) being less likely to receive vaccine.
+Finally, former drinkers (OR = 1.059, 95% CI = 1.010-1.111, p = 0.0172)
+were more likely to receive the vaccine compared to nondrinkers, and
+this was statistically significant. However, current drinkers (OR =
+0.978, 95% CI = 0.941 - 1.016) were less likely to receive vaccine
+compared to nondrinkers. Although this association was not statistically
+significant.
 
@@
 
-The results of our analysis indicate that several demographic and health-related factors are significantly associated with flu vaccine uptake.
-Our findings suggest that age, sex, race, education level, family income, and health status are all important predictors of vaccine uptake.
-Specifically, older individuals, females, individuals with higher education levels, and those with higher family incomes were more likely to have received the vaccine.
-On the other hand, individuals who identified as black or Hispanic were less likely to have received the vaccine, while Native Americans and Asians were more likely to have received it.
-In terms of health-related variables, individuals with asthma, cancer, and coronary heart disease were more likely to have received the vaccine compared to those without these conditions.
-Finally, our findings also suggest that smoking and drinking status are associated with vaccine uptake, with former smokers and drinkers more likely to have received the vaccine compared to nonsmokers and nondrinkers, respectively.
-These results have important implications for public health efforts aimed at increasing vaccine uptake, particularly among underserved and marginalized populations.
+The results of our analysis indicate that several demographic and
+health-related factors are significantly associated with flu vaccine
+uptake. Our findings suggest that age, sex, race, education level,
+family income, and health status are all important predictors of vaccine
+uptake. Specifically, older individuals, females, individuals with
+higher education levels, and those with higher family incomes were more
+likely to have received the vaccine. On the other hand, individuals who
+identified as black or Hispanic were less likely to have received the
+vaccine, while Native Americans and Asians were more likely to have
+received it. In terms of health-related variables, individuals with
+asthma, cancer, and coronary heart disease were more likely to have
+received the vaccine compared to those without these conditions.
+Finally, our findings also suggest that smoking and drinking status are
+associated with vaccine uptake, with former smokers and drinkers more
+likely to have received the vaccine compared to nonsmokers and
+nondrinkers, respectively. These results have important implications for
+public health efforts aimed at increasing vaccine uptake, particularly
+among underserved and marginalized populations.
 
@@
 
-While caution should be exercised when interpreting the significance of associations in observational studies due to potential confounding and reverse causality, this analysis provides valuable insights into identifying groups at risk for low influenza vaccine uptake over time.
-Limited studies have explored the relationship between a comprehensive list of factors and influenza vaccine coverage, and the large study population in this research improves statistical power, generalizability, precision, and accuracy of the findings.
-The design of the NHIS survey also allows for representative sampling of households, reducing sampling bias.
-Additionally, the survey questions were tailored to the month of each interview, enabling more detailed assessment of the seasonality of influenza vaccine schedule in the United States.
+While caution should be exercised when interpreting the significance of
+associations in observational studies due to potential confounding and
+reverse causality, this analysis provides valuable insights into
+identifying groups at risk for low influenza vaccine uptake over time.
+Limited studies have explored the relationship between a comprehensive
+list of factors and influenza vaccine coverage, and the large study
+population in this research improves statistical power,
+generalizability, precision, and accuracy of the findings. The design of
+the NHIS survey also allows for representative sampling of households,
+reducing sampling bias. Additionally, the survey questions were tailored
+to the month of each interview, enabling more detailed assessment of the
+seasonality of influenza vaccine schedule in the United States.
 
-In terms of limitations, as noted above, observational studies cannot demonstrate causality.
-The reason for not receiving an influenza vaccine is not an NHIS variable, which could have provided more insight.
-It is also important to note that the primary outcome of vaccination within the past 12 months was self-reported, and there was no way to verify using medical records, meaning that there is a risk of reporting bias due to recall, social desirability, or other factors.
-The reliability of self-report compared to health records for influenza vaccine is not known.
-Lastly, It is also unclear to what extent the COVID-19 pandemic and antivaccine movements have impacted vaccine behavior and hesitancy, which could be explored in future qualitative studies.
+In terms of limitations, as noted above, observational studies cannot
+demonstrate causality. The reason for not receiving an influenza vaccine
+is not an NHIS variable, which could have provided more insight. It is
+also important to note that the primary outcome of vaccination within
+the past 12 months was self-reported, and there was no way to verify
+using medical records, meaning that there is a risk of reporting bias
+due to recall, social desirability, or other factors. The reliability of
+self-report compared to health records for influenza vaccine is not
+known. Lastly, It is also unclear to what extent the COVID-19 pandemic
+and antivaccine movements have impacted vaccine behavior and hesitancy,
+which could be explored in future qualitative studies.
 
@@
 
-In conclusion, this study highlights the various demographic and health-related factors that are significantly associated with flu vaccine uptake.
-Understanding these predictors can help inform public health efforts aimed at increasing vaccine uptake, especially among marginalized populations.
-The findings suggest that targeted interventions may be necessary to address disparities in vaccine uptake among different racial and ethnic groups, as well as those with lower education levels and family incomes.
-Additionally, our findings on the association between smoking and drinking status and vaccine uptake may suggest the need for tailored messaging to encourage vaccination among these populations.
-Overall, these results can help guide public health interventions to promote vaccine uptake and ultimately reduce the burden of flu-related illness and hospitalizations.
+In conclusion, this study highlights the various demographic and
+health-related factors that are significantly associated with flu
+vaccine uptake. Understanding these predictors can help inform public
+health efforts aimed at increasing vaccine uptake, especially among
+marginalized populations. The findings suggest that targeted
+interventions may be necessary to address disparities in vaccine uptake
+among different racial and ethnic groups, as well as those with lower
+education levels and family incomes. Additionally, our findings on the
+association between smoking and drinking status and vaccine uptake may
+suggest the need for tailored messaging to encourage vaccination among
+these populations. Overall, these results can help guide public health
+interventions to promote vaccine uptake and ultimately reduce the burden
+of flu-related illness and hospitalizations.
 
